### 2019-04-18 code/rmarkdown/Supporting_information.rmd
first commit

---
title: "Supporting Information: Phylogeographic inference with microsatellites and SNPs"
author: |
  | Miguel Camacho-Sanchez${\ast^1}$, Guillermo Velo-Antón$^{1}$, Ana Veríssimo$^1$, Silvia Carvalho$^2$
  | $^1$ CIBIO/InBIO, Centro de Investigação em Biodiversidade e Recursos Genéticos da Universidade do Porto, Vairão, Portugal
  | Correspondence should be addressed to miguelcamachosachez@gmail.com
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  rmarkdown::pdf_document:
    toc: false
    keep_tex: true
    fig_caption: yes
    includes:
      in_header: preamble.tex
fontsize: 11pt
documentclass: article
subparagraph: true
---

### 2019-04-18 code/rmarkdown/article.Rmd
first commit

---
title: "Phylogeographic inference with microtellite and SNPs"
author: |
  | Miguel Camacho-Sanchez${\ast^1}$, Guillermo Velo-Antón$^{1}$, Ana Veríssimo$^1$, Silvia Carvalho$^2$
  | $^1$ CIBIO/InBIO, Centro de Investigação em Biodiversidade e Recursos Genéticos da Universidade do Porto, Vairão, Portugal
  | Correspondence should be addressed to miguelcamachosachez@gmail.com
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  rmarkdown::pdf_document:
    toc: false
    keep_tex: true
    fig_caption: yes
    includes:
      in_header: preamble.tex
fontsize: 11pt
documentclass: article
subparagraph: true
---

```{r, include = FALSE}
knitr::opts_knit$set(root.dir = normalizePath("../.."))
```

```{r, include = FALSE}
# restore session
gen <- readRDS("data/intermediate/raw_genotypes.RDS")
# load packages
# define helper functions
# load data

```

# MANUSCRIPT INFORMATION

\clearpage

# ABSTRACT

TODO

\clearpage

# 1 | INTRODUCTION

# 2 | MATERIALS AND METHODS

## Study area and species

## Data

# 3 | RESULTS

* result 1: sequencing
We sequenced a total of ´r adegenet::nInd(gen)´ samples which included 1 replica. DArTseq genotyped ´r adegenet::nLoc(gen)´SNPs in ´r nlevels(as.factor(gen@other$loc.metrics$CloneID))´ loci. The overall missing data was ´r ´ ()

* result 2: structure vs dapc

* result 3: genetic diversity

# 4 | DISCUSSION

* Overview

* Key result 1

* Key result 2

* Key result 3

* Concluding paragraph

# AUTHORS' CONTRIBUTIONS

TODO

# DATA ACCESSIBILITY

# ORCID

Miguel Camacho-Sanchez: 0000-0002-6385-7963 \newline
Guillermo Velo-Antón: 0000-0002-9483-5695 \newline
Ana Veríssimo: 0000-0003-3396-9822 \newline
JEFF ADAM PEDRO?
Silvia B. Carvalho: 0000-0003-4368-4708 \newline

# REFERENCES

### 2019-04-18 code/1_read_data.R

@@ -1,0 +1,49 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# April 2019
+###.............................................................................
+#GOAL: read dartseq and usat genotypes and metadata and save them as R objects
+#DESCRIPTION:
+#PROJECT: usat_snp (https://github.com/csmiguel/rana_iberica)
+###.............................................................................
+#  REQUIRED FILES:
+#   Description: read data/raw/readme
+#   Inpath: data/raw/
+#  OUTPUT:
+#    Description: raw genotypes plus metadata for dartseq samples
+#    Outpath: data/intermediate/raw_genotypes.rds & metadata_dartseq.rds
+#  DEPENDENCIES: functions/read_metadata.r & read_usats.r
+###.............................................................................
+
+library(dartR)
+library(dplyr)
+
+#1. DArTseq
+#path to singlerow dartseq SNPs genotypes
+path_2_genotypes <- c(
+  "data/raw/Report_DHyl17-2984_1_moreOrders_SNP_singlerow_2.csv",
+  "data/raw/Report_DPelo19-4013_SNP_mapping_2.csv")
+
+#read singlerow dartseq file into genlight object.
+gen_dart <- lapply(path_2_genotypes, dartR::gl.read.dart)
+names(gen_dart) <- c("hyla", "pelo")
+#read metadata
+source("code/functions/read_metadata.r")
+gen_dart %<>% metadata_Dartseq_Hyla() %>% metadata_Dartseq_Pelobates()
+#2. microsatellite
+source("code/functions/read_usats.r")
+  #genlight objects with metadata
+hyla_usat <- read_Hyla_molleri()
+pelo_usat <- read_Pelobates_cultripes()
+
+#create objects for exporting
+gen <- list(dart_hyla = gen_dart[[1]], dart_pelo = gen_dart[[2]],
+   usat_hyla = hyla_usat, usat_pelo = pelo_usat)
+metadata_dartseq <- list(metadata_hyla = metadata_hyla,
+  metadata_pelo = metadata_hyla)
+#note: usat metadata is already stored inside genind objects.
+
+saveRDS(gen, file = "data/intermediate/raw_genotypes.rds")
+saveRDS(metadata_dartseq, file = "data/intermediate/metadata_dartseq.rds")

### 2019-04-18 code/functions/gl_stats.r

@@ -1,0 +1,12 @@
+gl.stats <- function(genlight){
+  x <- genlight
+  firstc <- which(names(x@other$loc.metrics) == "CallRate")
+  lastc <- ncol(x@other$loc.metrics)
+  metrics <<- x@other$loc.metrics[, c(firstc:lastc)]
+  mean1 <- apply(metrics, 2, mean)
+  q1 <- apply(metrics, 2, quantile)
+  sd1 <- apply(metrics, 2, sd)
+  all1 <- cbind(mean1, t(q1), sd1)
+  colnames(all1) <- c("mean", dimnames(q1)[[1]], "sd")
+  return(all1)
+}

### 2019-04-18 code/functions/ind_miss.r

@@ -1,0 +1,7 @@
+#calculate missingness per individual from genlight object
+ind_miss <- function(gl){
+  apply(as.matrix(gl), 1, function(x){
+  sum(is.na(x))
+  }
+  ) / adegenet::nLoc(gl)
+}

### 2019-04-18 code/functions/read_metadata.r

@@ -1,0 +1,43 @@
+#read metadata hyla
+metadata_Dartseq_Hyla <- function(x){
+  f <- "data/raw/metadata_hyla_dartseq.xlsx"
+  m <- readxl::read_excel(f)
+  m_darseq_id <- m[["dartseqID"]]
+  assertthat::assert_that(class(x$hyla) == "genlight")
+  g_dartseq_id <- indNames(x$hyla) %>% gsub(" ", "", .)
+assertthat::assert_that(!is.null(c(m[["dartseqID"]],
+      m[["locality"]],
+      m[["id"]])))
+  #fix sample names
+  indNames(x$hyla) <- sapply(g_dartseq_id, function(x){
+    which(m[["dartseqID"]] == x) %>% m[["id"]][.]
+  })
+  #fix pop names
+  pop(x$hyla) <- sapply(g_dartseq_id, function(x){
+    which(m[["dartseqID"]] == x) %>% m[["locality"]][.]
+  })
+  metadata_hyla <<- m
+  return(x)
+}
+
+metadata_Dartseq_Pelobates <- function(x){
+  f <- "data/raw/2019-01-03_meta_Pelobates.csv"
+  m <- read.csv(f)
+  m %<>% mutate(locality = gsub(pattern = ",.*$", replacement = "", site)) %>%
+    rename(full_locality = site) %>%
+    rename(dartseqID = vaucher) %>%
+    rename(latitude = lat) %>%
+    rename(longitude = lon)
+
+  assertthat::assert_that(class(x$pelo) == "genlight")
+  m_darseq_id <- m[["dartseqID"]]
+  g_dartseq_id <- indNames(x$pelo)
+  assertthat::assert_that(all(g_dartseq_id %in% m_darseq_id))
+  #fix pop names
+  assertthat::assert_that(!is.null(c(m[["dartseqID"]], m[["locality"]])))
+  pop(x$pelo) <- sapply(g_dartseq_id, function(x){
+    which(m[["dartseqID"]] == x) %>% m[["locality"]][.]
+  })
+  metadata_pelo <<- m
+  return(x)
+}

### 2019-04-18 code/functions/read_usats.r

@@ -1,0 +1,132 @@
+#this piece of code has been copied from Jeff's. I have modified at likes marked
+#with #m
+read_Pelobates_cultripes <- function() {
+  ## read data
+  f <- "data/raw/PcultripesPhylogeography.gen" #m
+  o <- adegenet::read.genepop(f, ncode = 3)
+  s <-
+    read.table("data/raw/table-s1-data.csv",
+               header = TRUE, sep = ",", stringsAsFactors = FALSE) %>%
+    tibble::as_tibble() %>%
+    dplyr::filter(N > 0) %>%
+    plyr::ddply("ID", function(x) {
+      # iterate over codes and interpolate codes specified as ranges
+      plyr::ldply(
+        stringr::str_squish(strsplit(x$Sample.Codes, ",")[[1]]),
+        function(y) {
+          # extract codes
+          if (grepl("-", y, fixed = TRUE)) {
+            ## find start and end codes
+            start <- strsplit(y, "-", fixed = TRUE)[[1]]
+            end <- start[[2]]
+            start <- start[[1]]
+            ## find character prefix for code
+            prefix <- gsub("[[:digit:]]", "", start)
+            ## interpolate numbers
+            start_number <- as.numeric(gsub(prefix, "", start, fixed = TRUE))
+            end_number <- as.numeric(gsub(prefix, "", end, fixed = TRUE))
+            assertthat::assert_that(is.finite(start_number),
+                                    is.finite(end_number))
+            ## create codes
+            codes <- paste0(prefix, seq(start_number, end_number))
+          } else {
+            codes <- y
+          }
+          # return table with sample codes per locality
+          data.frame(ID = x$ID, Locality = x$Locality, Sample.Code = codes,
+                     Longitude = x$Longitude, Latitude = x$Latitude,
+                     stringsAsFactors = FALSE)
+        })
+    }) %>%
+    tibble::as_tibble()
+  ## add in population data,
+  ### since we only have sample ids for individuals with mtDNA data,
+  ### we don't have complete information on which samples belong to which
+  ### populations, but if we known that sample A is in the same population as
+  ### as sample B, and we known that sample B is in population A then we know
[…]

### 2019-04-18 code/parameters/gl_cleaning.r

@@ -1,0 +1,7 @@
+ind_miss_thresh <- 0.65
+repavg_threshold <- 0.98
+locus_callrate_threshold <- 0.8
+sec_method <- "best"
+min_maf <- 0.04
+hwe_alpha <- 0.05
+rep_sample <- "7865"

### 2019-04-18 code/1_read_data.R
dart snp cleaning done

@@ -6,7 +6,7 @@
 ###.............................................................................
 #GOAL: read dartseq and usat genotypes and metadata and save them as R objects
 #DESCRIPTION:
-#PROJECT: usat_snp (https://github.com/csmiguel/rana_iberica)
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
 ###.............................................................................
 #  REQUIRED FILES:
 #   Description: read data/raw/readme

### 2019-04-18 code/2_genotype_filtering.R

@@ -1,0 +1,68 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# April 2019
+###.............................................................................
+#GOAL: filter DARrTseq genotypes from Hyla and Pelobates
+#DESCRIPTION:
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+#  REQUIRED FILES:
+#   Description: genlight objects with raw genotypes for Hyla and Pelobates
+#   Inpath: data/intermediate/raw_genotypes.rds
+#  OUTPUT:
+#    Description: raw genotypes in genlight format
+#    Outpath: data/intermediate/filtered_genotypes.rds
+###.............................................................................
+library(dartR)
+library(dplyr)
+
+source("code/functions/gl_stats.r")
+source("code/functions/ind_miss.r")
+
+input_path <- paste0("data/intermediate/raw_genotypes.rds")
+gen <- readRDS(file = input_path)
+
+#summary statistics for raw data
+raw_stats <- geno_stats(genlist = gen, data = "raw")
+##################
+#Filtering
+source("code/parameters/gl_cleaning.r")
+sink(file = "data/intermediate/filtering.log")
+gen_filt <- grep("dart", names(gen)) %>% #only for dart genotypes
+  {genfiltnames <<- names(gen)[.]; .} %>%
+  seq_along() %>%
+  lapply(function(i){
+    sp <- paste(names(gen)[[i]])
+    cat(paste0("#log for ", sp), "\n")
+    dartR::gl.filter.callrate(gen[[i]], method = "ind",
+    threshold = 1 - ind_miss_thresh, v = 5) %>%
+    dartR::gl.filter.repavg(repavg_threshold, v = 5) %>%
+    dartR::gl.filter.callrate(method = "loc",
+      threshold = locus_callrate_threshold, v = 5, recalc = T) %>%
+    dartR::gl.filter.secondaries(method = sec_method, v = 5) %>%
+    dartR::gl.filter.maf(threshold = min_maf, v = 5) %>%
+    {real_pop <<- pop(.); .} %>%
[…]

### 2019-04-18 code/functions/gl_stats.r

@@ -1,12 +1,32 @@
-gl.stats <- function(genlight){
-  x <- genlight
+loci.stats <- function(x){
   firstc <- which(names(x@other$loc.metrics) == "CallRate")
-  lastc <- ncol(x@other$loc.metrics)
-  metrics <<- x@other$loc.metrics[, c(firstc:lastc)]
+  lastc <- which(names(x@other$loc.metrics) == "RepAvg")
+  metrics <- x@other$loc.metrics[, c(firstc:lastc)]
   mean1 <- apply(metrics, 2, mean)
   q1 <- apply(metrics, 2, quantile)
   sd1 <- apply(metrics, 2, sd)
   all1 <- cbind(mean1, t(q1), sd1)
   colnames(all1) <- c("mean", dimnames(q1)[[1]], "sd")
   return(all1)
+}
+
+geno_stats <- function(pattern = "dart", genlist, data = c("raw", "filt")){
+  grep(pattern, names(genlist)) %>% #only for dart genotypes
+  seq_along() %>%
+  lapply(function(i){
+    sp <- paste(names(genlist)[[i]])
+    summary_metrics <- loci.stats(genlist[[i]]) #metrics per loci
+    ind_miss_raw <- ind_miss(genlist[[i]]) #missingness per individual
+      #plot individual missingness
+    pdf(file = paste0("data/intermediate/", data, "_", sp,
+    "_ind_missingness.pdf"), width = 6, height = 4)
+          plot(sort(ind_miss_raw),
+            main = paste0("Sorted per indiviudal missingness in ", data,
+            " data from ", sp),
+            ylab = "Proportion of missing loci")
+    dev.off()
+    list(genotypes = paste0(data, "_", sp), summary_metrics = summary_metrics,
+      ind_miss_raw = ind_miss_raw)
+    }
+  )
 }

### 2019-04-18 code/parameters/gl_cleaning.r

@@ -2,6 +2,5 @@ ind_miss_thresh <- 0.65
 repavg_threshold <- 0.98
 locus_callrate_threshold <- 0.8
 sec_method <- "best"
-min_maf <- 0.04
+min_maf <- 0.02 #eg: 95samples*2ploidy*0.8callrate*0.02maf = 3 alleles
 hwe_alpha <- 0.05
-rep_sample <- "7865"

### 2019-05-09 code/rmarkdown/Figures.rmd
start methods

---
title: "Figures"
fontsize: 11pt
documentclass: article
subparagraph: true
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: no
    includes:
      in_header: preamble.tex
---

```{r, include = FALSE}
knitr::opts_knit$set(root.dir = normalizePath("../.."))
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dev = "png", dev.args = list(type = "cairo"), dpi = 500, fig.align = "center")
```

##Figure 1:
```{r, fig.height = 4.5, fig.width = 4.5, fig.cap = "Distribution and populations sampled in *Hyla molleri* (panel A) and *Pelobates cultripes* (panel B), with corresponding genotype dataset: microsatellite (triangles)  and SNP (circles). The number inside each symbol refer to the population ID for microsatellite data in Sánchez-Montes et al. (2019)(*H. molleri), Gutiérrez-Rodríguez et al. (2017)(*P. cultripes) and SNPs Table 1. The sample sizes for each population are indicated in parenthesis. The distribution for each species is indicated in shaded area, downloaded from the IUCN on May 2019.(xx note: because H. molleri is not in the IUCN database still, I used the distribution of H. arborea)"}

```

### 2019-05-09 code/00_raster_iberianPeninsula.R

@@ -1,0 +1,22 @@
+#create raster file for the Iberian Peninsula
+path <- "data/raw/raster_iberian_peninsula.rds"
+assertthat::assert_that(!file.exists(path),
+    msg = paste(path, "already exists"))
+
+mid_long <- c(-7.5, -2.5, 2.5)
+mid_lat <- c(37.5, 42.5)
+mid_latlon <- expand.grid(mid_lat, mid_long)
+
+o <- as.list(numeric(nrow(mid_latlon)))
+#use SRTM 90 with 90 m resolution at the equator.
+#for tiles see:http://srtm.csi.cgiar.org/srtmdata/
+for (i in seq_along(o)){
+  o[[i]] <- raster::getData("SRTM", download = T, "data/intermediate",
+                            lat = mid_latlon[i, 1], lon = mid_latlon[i, 2]) %>%
+    raster::aggregate(fact = 3)
+}
+
+o$fun <- mean
+o$na.rm <- TRUE
+mapr <- do.call(raster::mosaic, o)
+saveRDS(mapr, path)

### 2019-05-09 code/01_read_data.R

@@ -1,0 +1,49 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# April 2019
+###.............................................................................
+#GOAL: read dartseq and usat genotypes and metadata and save them as R objects
+#DESCRIPTION:
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+#  REQUIRED FILES:
+#   Description: read data/raw/readme
+#   Inpath: data/raw/
+#  OUTPUT:
+#    Description: raw genotypes plus metadata for dartseq samples
+#    Outpath: data/intermediate/raw_genotypes.rds & metadata_dartseq.rds
+#  DEPENDENCIES: functions/read_metadata.r & read_usats.r
+###.............................................................................
+
+library(dartR)
+library(dplyr)
+
+#1. DArTseq
+#path to singlerow dartseq SNPs genotypes
+path_2_genotypes <- c(
+  "data/raw/Report_DHyl17-2984_1_moreOrders_SNP_singlerow_2.csv",
+  "data/raw/Report_DPelo19-4013_SNP_mapping_2.csv")
+
+#read singlerow dartseq file into genlight object.
+gen_dart <- lapply(path_2_genotypes, dartR::gl.read.dart)
+names(gen_dart) <- c("hyla", "pelo")
+#read metadata
+source("code/functions/read_metadata.r")
+gen_dart %<>% metadata_Dartseq_Hyla() %>% metadata_Dartseq_Pelobates()
+#2. microsatellite
+source("code/functions/read_usats.r")
+  #genlight objects with metadata
+hyla_usat <- read_Hyla_molleri()
+pelo_usat <- read_Pelobates_cultripes()
+
+#create objects for exporting
+gen <- list(dart_hyla = gen_dart[[1]], dart_pelo = gen_dart[[2]],
+   usat_hyla = hyla_usat, usat_pelo = pelo_usat)
+metadata_dartseq <- list(metadata_hyla = metadata_hyla,
+  metadata_pelo = metadata_hyla)
+#note: usat metadata is already stored inside genind objects.
+
+saveRDS(gen, file = "data/intermediate/raw_genotypes.rds")
+saveRDS(metadata_dartseq, file = "data/intermediate/metadata_dartseq.rds")

### 2019-05-09 code/02_genotype_filtering.R

@@ -1,0 +1,75 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# April 2019
+###.............................................................................
+#GOAL: filter DARrTseq genotypes from Hyla and Pelobates
+#DESCRIPTION:
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+#  REQUIRED FILES:
+#   Description: genlight objects with raw genotypes for Hyla and Pelobates
+#   Inpath: data/intermediate/raw_genotypes.rds
+#  OUTPUT:
+#    Description: raw genotypes in genlight format
+#    Outpath: data/intermediate/filtered_genotypes.rds
+###.............................................................................
+library(dartR)
+library(dplyr)
+
+source("code/functions/gl_stats.r")
+source("code/functions/ind_miss.r")
+source("code/functions/remove_replica.r")
+
+input_path <- paste0("data/intermediate/raw_genotypes.rds")
+gen <- readRDS(file = input_path)
+
+#summary statistics for raw data
+raw_stats <- geno_stats(genlist = gen, data = "raw")
+
+##################
+#Filtering
+sink(file = "data/intermediate/filtering.log")
+  #remove replica from Pelobates:
+      #for Pelobates Ana included a blind
+      #replica named 108541 from sample 106854.
+gen$dart_pelo <- remove_replica(gen = gen$dart_pelo, "108541", "106854")
+  #filter SNPs
+source("code/parameters/gl_cleaning.r")
+gen_filt <- grep("dart", names(gen)) %>% #only for dart genotypes
+  {genfiltnames <<- names(gen)[.]; .} %>%
+  seq_along() %>%
+  lapply(function(i){
+    sp <- paste(names(gen)[[i]])
+    cat(paste0("#log for ", sp), "\n")
+    dartR::gl.filter.callrate(gen[[i]], method = "ind",
+    threshold = 1 - ind_miss_thresh, v = 5) %>%
+    dartR::gl.filter.repavg(repavg_threshold, v = 5) %>%
+    dartR::gl.filter.callrate(method = "loc",
[…]

### 2019-05-09 code/08_Read_Structure.R

@@ -1,0 +1,5 @@
+path <- "/Users/miguelcamacho/Dropbox/Research/CiBIO/NGC/dartseq_genotyping/structure/dartseq/run1"
+
+read str stlog  cat K6_rep4.stlog | grep [00]: > lk.txt
+h <- read.table("data/intermediate/lk.txt", skip =1)
+plot(h$X0.068, type = "l")

### 2019-05-09 code/6_Prepare_structure.R

@@ -1,0 +1,68 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# April 2019
+###.............................................................................
+#GOAL: export data into structure format
+#DESCRIPTION: export data into structure format and create params files
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+#  REQUIRED FILES:
+#   Description:
+#   Inpath:
+#  OUTPUT:
+#    Description: raw genotypes plus metadata for dartseq samples
+#    Outpath:
+#  DEPENDENCIES:
+###.............................................................................
+library(dartR)
+library(dplyr)
+
+#1. Create STRUCTURE input files from dart and usat genotypes
+input_path <- paste0("data/intermediate/filt_genotypes.rds")
+gen <- readRDS(file = input_path)
+source("code/functions/edit_params.r")
+
+#warning dartR::gi2gl %>% gl2structure does not format format properly
+#structure files. But, hierfstat::write.struct has a bug so that it does not
+#print individual names in the ilab argument. I overcame this by modifiying the
+#dataframe with genotypes:
+names(gen) %>% seq_along() %>%
+  sapply(function(x){
+    dir_str <- file.path("data/intermediate", paste0("str_", names(gen)[x]))
+    dir.create(dir_str, showWarnings = T)
+    str_file <- file.path(dir_str, paste0(names(gen)[x], ".str"))
+    edit_mainparams(gl = gen[[x]], name = names(gen)[x],
+      str_file = str_file, dir_str = dir_str)
+    edit_extraparams(gl = gen[[x]], name = names(gen)[x],
+      str_file = str_file, dir_str = dir_str)
+    if (class(gen[[x]]) == "genind"){
+      hierfstat::genind2hierfstat(gen[[x]]) %>%
+      dplyr::mutate(pop = as.factor(indNames(gen[[x]]))) %>%
+      hierfstat::write.struct(fname = str_file)
+      } else if (class(gen[[x]]) == "genlight") {
+      dartR::gl2structure(gen[[x]], outfile = str_file)
+    }
+  })
+
[…]

### 2019-05-09 code/functions/edit_params.r

@@ -1,0 +1,41 @@
+source("code/parameters/structure.r")
+
+#edit mainparams file for each STRUCTURE run
+
+edit_mainparams <- function(gl = gl, name = name,
+  str_file = str_file, dir_str = dir_str){
+  param <- readLines("data/raw/mainparams")
+  param[16] <- sub("xx",  10, param[16])#max pops (K)
+  param[27] <- sub("xx",  adegenet::nInd(gl), param[27])#number of individual
+  param[28] <- sub("xx",  adegenet::nLoc(gl), param[28])#number of loci
+  param[22] <- sub("xx",  str_file, param[22])#infile.
+  #structure_threader will take output and input infiles from -o and -i, but
+  #not from the params files
+    output <- file.path("data/final", paste0("str_", name, "_run"))
+  param[23] <- sub("xx",  output, param[23])#outfile
+  if (grepl("dart*", name) | is.numeric(name)){
+    param[17] <- sub("xx", burnin_snp, param[17])#burnin length
+    param[18] <- sub("xx", run_snp, param[18])#run length
+  } else if (grepl("usat*", name)){
+    param[17] <- sub("xx", burnin_usat, param[17])#burnin length
+    param[18] <- sub("xx", run_usat, param[18])#run length
+  }
+  assertthat::assert_that(all(grepl("xx", param) == F),
+    msg = "wrong replacement in mainparams file")
+  mainparams_path <- file.path(dir_str, "mainparams")
+  writeLines(param, mainparams_path)
+}
+
+edit_extraparams <- function(gl = gl, name = name,
+  str_file = str_file, dir_str = dir_str){
+  param <- readLines("data/raw/extraparams")
+  if (grepl("dart*", name) | is.numeric(name)){
+    param[93] <- sub("xx", run_snp / 200, param[93])#burnin length
+  } else if (grepl("usat*", name)){
+    param[93] <- sub("xx", run_usat / 200, param[93])#burnin length
+  }
+  assertthat::assert_that(all(grepl("xx", param) == F),
+    msg = "wrong replacement in extraparams file")
+  mainparams_path <- file.path(dir_str, "extraparams")
+  writeLines(param, mainparams_path)
+}

### 2019-05-09 code/functions/map_distribution.r

@@ -1,0 +1,40 @@
+#create map with the distribution of the samples and native distribution of the species
+#read raster and shape files
+p_raster <- "data/raw/raster_iberian_peninsula.rds"
+mapr <- readRDS(p_raster)
+  #IUCN distribution
+p_pelo <- "data/raw/redlist_species_data_44f009a9-b3b2-4608-a818-24c3dd714900"
+p_hyla <- "data/raw/redlist_species_data_eff4cf11-63bb-4131-bba3-104c3358697b"
+shp_pelo <- rgdal::readOGR(dsn = p_pelo, layer = "data_0") %>%
+  sp::spTransform(mapr@crs)
+shp_hyla <- rgdal::readOGR(dsn = p_hyla, layer = "data_0") %>%
+  sp::spTransform(mapr@crs)
+
+  #dartseq locations
+  p_dartmeta <- ""
+  dartmeta <- readRDS()
+  #microsatellite locations
+  p_gen <- ""
+  gen <- readRDS()
+
+assertthat::assert_that(as.character(shp_hyla@data$BINOMIAL)[1] == "Hyla arborea")
+assertthat::assert_that(as.character(shp_pelo@data$BINOMIAL)[1] == "Pelobates cultripes")
+
+#Plot distribution
+plot(mapr, breaks = c(0, 500, 1000, 1500, 2000, 3500),
+      col = gray.colors(5, start = 0.9, end = 0.3)) #issue margins
+plot(shp_pelo, add = T, border = "black",
+  col = grey(0.5, alpha = 0.3))
+title(shp_pelo@data$BINOMIAL[1])
+
+#several options for plot:
+# plot(h, col = gray.colors(4))
+# plot(h, col = gray.colors(4, start = 0.9, end = 0.3))
+# plot(h, breaks = c(0, 500, 1000, 1500, 2000, 3500),
+#      col = c("lightgreen", "darkgreen", "yellow", "brown", "purple"))
+# plot(h, breaks = c(0, 500, 1000, 1500, 2000, 3500),
+#      col = gray.colors(5, start = 0.9, end = 0.3),
+#      xlim = c(-10, 5))
+# spplot(mapr, maxpixels= 500000,
+#     col.regions = gray.colors(5, start = 0.9, end = 0.3),
+#     at = c(0, 500, 1000, 1500, 2000, 3500))

### 2019-05-09 code/parameters/structure.r

@@ -1,4 +1,4 @@
 burnin_usat = 500000 #as in Gutiérrez-Rodríguez et al. 2017 and Sánchez-Montes et al. 2019
-burnin_snp = 10000
+burnin_snp = 5000
 run_usat = 1000000 #as in Gutiérrez-Rodríguez et al. 2017 and Sánchez-Montes et al. 2019
-run_snp = 20000
+run_snp = 10000

### 2019-05-09 code/rmarkdown/article.Rmd
cleaning all files

@@
 ```
+# Manuscript information
 
-# MANUSCRIPT INFORMATION
+Format: add journal title  \newline
+Abstract word count: xx \newline
+Main word count: xx  \newline
+Number of references: xx  \newline
+Number of figures: xx  \newline
+Number of tables: xx  \newline
+Number of text boxes: xx  \newline
+Number of equations: xx  \newline
+Keywords: microtellites, DArTseq, SNPs, neighbour joining, structure, phylogeography, Hyla molleri, Pelobates cultripes, Iberian Peninsula   \newline
+Spelling: British English (en-GB)   \newline
+\clearpage
+
+# Abstract
 
@@
 
-# ABSTRACT
+# 1. Introduction
+**general overview about the usat vs snp**
+1-which are the characteristics of each marker
+2-how informative are they
+3-pros and cons for both
+4-limitations
+5-previous studies focusing on usats vs SNPs
+**our case study**
+1-study species: characteristics, distrigution, why them
+2-markers datasets
+**brief methodology with results and conclusions**
 
-TODO
-
-\clearpage
-
-# 1 | INTRODUCTION
-
-# 2 | MATERIALS AND METHODS
+# 2. Materials and Methods
+## 2.1 Data collection
+Our analysis are built upon microsatellite and SNP datasets for two Iberian amphibians: *Hyla molleri* and *Pelobates cultripes*.
+For each species, the microtellite and SNP datasets come from different sample sets but covering both most of their native range (Figure 1).
+The microsatellite genotypes were taken from already-published studies.
+For *Hyla molleri*, we gathered the microsatellite genotypes from *Appendix 1* in Sánchez-Montes et al.
+(2019).
+They consisted of `r adegenet::nInd(gen$usat_hyla)` samples with `r adegenet::nLoc(gen$usat_hyla)` genotyped microsatellites.
+For *Pelobates cultripes*, we requested the genotypes to the authors in Gutiérrez-Rodríguez et al.
+(2017).
+This dataset consisted of `r adegenet::nInd(gen$usat_pelo)` samples with `r adegenet::nLoc(gen$usat_pelo)` genotyped microsatellites.
+The SNP data was generated using DArTseq technology (xx).
+##
 
@@
 Ana Veríssimo: 0000-0003-3396-9822 \newline
-JEFF ADAM PEDRO?
+Jeffrey O. Hanson: 0000-0002-4716-6134
+Adam Marques: 0000-0002-6070-0506
 Silvia B. Carvalho: 0000-0003-4368-4708 \newline
 
+\clearpage
+
 # REFERENCES

### 2019-05-09 code/1_read_data.R

@@ -1,49 +1,0 @@
-###.............................................................................
-# (c) Miguel Camacho Sánchez
-# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
-# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
-# April 2019
-###.............................................................................
-#GOAL: read dartseq and usat genotypes and metadata and save them as R objects
-#DESCRIPTION:
-#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
-###.............................................................................
-#  REQUIRED FILES:
-#   Description: read data/raw/readme
-#   Inpath: data/raw/
-#  OUTPUT:
-#    Description: raw genotypes plus metadata for dartseq samples
-#    Outpath: data/intermediate/raw_genotypes.rds & metadata_dartseq.rds
-#  DEPENDENCIES: functions/read_metadata.r & read_usats.r
-###.............................................................................
-
-library(dartR)
-library(dplyr)
-
-#1. DArTseq
-#path to singlerow dartseq SNPs genotypes
-path_2_genotypes <- c(
-  "data/raw/Report_DHyl17-2984_1_moreOrders_SNP_singlerow_2.csv",
-  "data/raw/Report_DPelo19-4013_SNP_mapping_2.csv")
-
-#read singlerow dartseq file into genlight object.
-gen_dart <- lapply(path_2_genotypes, dartR::gl.read.dart)
-names(gen_dart) <- c("hyla", "pelo")
-#read metadata
-source("code/functions/read_metadata.r")
-gen_dart %<>% metadata_Dartseq_Hyla() %>% metadata_Dartseq_Pelobates()
-#2. microsatellite
-source("code/functions/read_usats.r")
-  #genlight objects with metadata
-hyla_usat <- read_Hyla_molleri()
-pelo_usat <- read_Pelobates_cultripes()
-
-#create objects for exporting
-gen <- list(dart_hyla = gen_dart[[1]], dart_pelo = gen_dart[[2]],
-   usat_hyla = hyla_usat, usat_pelo = pelo_usat)
-metadata_dartseq <- list(metadata_hyla = metadata_hyla,
-  metadata_pelo = metadata_hyla)
-#note: usat metadata is already stored inside genind objects.
-
-saveRDS(gen, file = "data/intermediate/raw_genotypes.rds")
-saveRDS(metadata_dartseq, file = "data/intermediate/metadata_dartseq.rds")

### 2019-05-09 code/2_genotype_filtering.R

@@ -1,75 +1,0 @@
-###.............................................................................
-# (c) Miguel Camacho Sánchez
-# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
-# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
-# April 2019
-###.............................................................................
-#GOAL: filter DARrTseq genotypes from Hyla and Pelobates
-#DESCRIPTION:
-#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
-###.............................................................................
-#  REQUIRED FILES:
-#   Description: genlight objects with raw genotypes for Hyla and Pelobates
-#   Inpath: data/intermediate/raw_genotypes.rds
-#  OUTPUT:
-#    Description: raw genotypes in genlight format
-#    Outpath: data/intermediate/filtered_genotypes.rds
-###.............................................................................
-library(dartR)
-library(dplyr)
-
-source("code/functions/gl_stats.r")
-source("code/functions/ind_miss.r")
-source("code/functions/remove_replica.r")
-
-input_path <- paste0("data/intermediate/raw_genotypes.rds")
-gen <- readRDS(file = input_path)
-
-#summary statistics for raw data
-raw_stats <- geno_stats(genlist = gen, data = "raw")
-
-##################
-#Filtering
-sink(file = "data/intermediate/filtering.log")
-  #remove replica from Pelobates:
-      #for Pelobates Ana included a blind
-      #replica named 108541 from sample 106854.
-gen$dart_pelo <- remove_replica(gen = gen$dart_pelo, "108541", "106854")
-  #filter SNPs
-source("code/parameters/gl_cleaning.r")
-gen_filt <- grep("dart", names(gen)) %>% #only for dart genotypes
-  {genfiltnames <<- names(gen)[.]; .} %>%
-  seq_along() %>%
-  lapply(function(i){
-    sp <- paste(names(gen)[[i]])
-    cat(paste0("#log for ", sp), "\n")
-    dartR::gl.filter.callrate(gen[[i]], method = "ind",
-    threshold = 1 - ind_miss_thresh, v = 5) %>%
-    dartR::gl.filter.repavg(repavg_threshold, v = 5) %>%
-    dartR::gl.filter.callrate(method = "loc",
[…]

### 2019-05-09 code/3_Phylogenetic_tree.R

@@ -1,56 +1,0 @@
-###.............................................................................
-# (c) Miguel Camacho Sánchez
-# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
-# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
-# April 2019
-###.............................................................................
-#GOAL: produce phylogenetic trees for microsatellite and snp data
-#DESCRIPTION:
-#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
-###.............................................................................
-#  REQUIRED FILES:
-#   Description:
-#   Inpath:
-#  OUTPUT:
-#    Description: raw genotypes plus metadata for dartseq samples
-#    Outpath:
-#  DEPENDENCIES:
-###.............................................................................
-library(adegenet)
-library(assertthat)
-
-input_path <- paste0("data/intermediate/filt_genotypes.rds")
-gen <- readRDS(file = input_path)
-
-source("code/functions/bootstrap_trees.r")
-source("code/functions/create_trees.r")
-
-nboot <- 100 #number of bootstrap
-
-tr <- names(gen) %>% seq_along() %>%
-sapply(function(j){
-  if (adegenet::nLoc(gen[[j]]) < 100) type <- "usat"
-    else if (adegenet::nLoc(gen[[j]]) > 100) type <- "snp"
-  sp <- names(gen[j])
-  temp <- list()
-  original_trees <- create_trees(gen[[j]], type = type)
-  boots_trees <- bootstrap_nj(gen[[j]], nboot, type = type)
-  temp[[1]] <- c(original_trees, boots_trees)
-  names(temp) <- sp
-  temp
-})
-
-
-#calculate bootstrap support
-bt <- names(tr) %>% seq_along() %>%
-sapply(function(x){
-  sp <- names(tr[x])
-  temp <- list()
-  temp[[1]] <- ape::countBipartitions(tr[[x]][[1]], tr[[x]][-1])
-  names(temp) <- sp
-  temp
-})
-
-#save trees (first tree is the real tree)
-saveRDS(tr, "data/intermediate/bt_trees.rds")
-saveRDS(bt, "data/intermediate/boot_support.rds") #bootstrap support

### 2019-05-09 code/3_bootstrap_tree.R

@@ -1,79 +1,0 @@
-###.............................................................................
-# (c) Miguel Camacho Sánchez
-# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
-# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
-# April 2019
-###.............................................................................
-#GOAL: produce phylogenetic trees for microsatellite and snp data
-#DESCRIPTION:
-#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
-###.............................................................................
-#  REQUIRED FILES:
-#   Description:
-#   Inpath:
-#  OUTPUT:
-#    Description: raw genotypes plus metadata for dartseq samples
-#    Outpath:
-#  DEPENDENCIES:
-###.............................................................................
-library(adegenet)
-library(assertthat)
-library(dplyr)
-library(ape)
-
-input_path <- paste0("data/intermediate/filt_genotypes.rds")
-gen <- readRDS(file = input_path)
-
-source("code/functions/bootstrap_trees.r")
-source("code/functions/create_trees.r")
-
-source("code/parameters/boot.r") #number of bootstrap
-
-tr <- names(gen) %>% seq_along() %>%
-sapply(function(j){
-  if (adegenet::nLoc(gen[[j]]) < 100) type <- "usat"
-    else if (adegenet::nLoc(gen[[j]]) > 100) type <- "snp"
-  sp <- names(gen[j])
-  temp <- list()
-  original_trees <- create_trees(gen[[j]], type = type)
-  boots_trees <- bootstrap_nj(gen[[j]], nboot, type = type)
-  temp[[1]] <- c(original_trees, boots_trees)
-  names(temp) <- sp
-  temp
-})
-
-
-#calculate bootstrap support
-bt <- names(tr) %>% seq_along() %>%
-sapply(function(x){
-  sp <- names(tr[x])
-  temp <- list()
-  phy <- reorder(tr[[x]][[1]], "postorder")
-  ints <- phy$edge[, 2] > ape::Ntip(phy)
-  ans <- ape::countBipartitions(phy, tr[[x]][-1])
-  temp[[1]] <- c(nboot, ans[order(phy$edge[ints, 2])])
-  names(temp) <- sp
-  temp
-})
-
-#plot trees
-  nodecol <- bt
-for (i in seq_along(names(bt))){
[…]

### 2019-05-09 code/4_BS_analysis.R

@@ -1,39 +1,0 @@
-###.............................................................................
-# (c) Miguel Camacho Sánchez
-# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
-# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
-# April 2019
-###.............................................................................
-#GOAL: analysis of bootstrap support
-#DESCRIPTION:
-#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
-###.............................................................................
-#  REQUIRED FILES:
-#   Description:
-#   Inpath:
-#  OUTPUT:
-#    Description: raw genotypes plus metadata for dartseq samples
-#    Outpath:
-#  DEPENDENCIES:
-###.............................................................................
-library(adegenet)
-library(assertthat)
-library(dplyr)
-library(ape)
-
-input_path1 <- paste0("data/intermediate/bt_trees.rds")
-input_path2 <- paste0("data/intermediate/boot_support.rds")
-tr <- readRDS(file = input_path1)
-bt <- readRDS(file = input_path2)
-
-source("code/functions/bs_analysis.r")
-source("code/parameters/boot.r")
-
-#plot bs analisis
-for (m in 1:3){
-pdf(file = paste0("data/intermediate/bs_analysis_mode", m, ".pdf"),
- height = 8, width = 10)
-par(mfrow = c(2, 2))
-for (p in 1:length(tr)) bs_analysis(tr[p], bt[[p]], mode = m)
-dev.off()
-}

### 2019-05-09 code/5_subsampling_snps.R

@@ -1,108 +1,0 @@
-###.............................................................................
-# (c) Miguel Camacho Sánchez
-# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
-# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
-# April 2019
-###.............................................................................
-#GOAL: export data into structure format
-#DESCRIPTION: export data into structure format and create params files
-#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
-###.............................................................................
-#  REQUIRED FILES:
-#   Description:
-#   Inpath:
-#  OUTPUT:
-#    Description: raw genotypes plus metadata for dartseq samples
-#    Outpath:
-#  DEPENDENCIES:
-###.............................................................................
-library(adegenet)
-library(assertthat)
-library(dplyr)
-library(ape)
-
-input_path1 <- paste0("data/intermediate/filt_genotypes.rds")
-gen <- readRDS(file = input_path1)
-input_path2 <- paste0("data/intermediate/bt_trees.rds")
-tr <- readRDS(file = input_path2)
-input_path3 <- paste0("data/intermediate/boot_support.rds")
-bt <- readRDS(file = input_path3)
-max_bt <- max(unlist(bt))
-bt <- sapply(function(x) x / max_bt)
-
-source("code/parameters/boot.r")
-source("code/functions/bootstrap_trees.r")
-source("code/functions/create_trees.r")
-
-#subset data
-pelo <- gen$dart_pelo
-original_trees <- lapply(tr, function(x) x[[1]])
-rm(tr)
-
-#create trees from subsampled dataset
-s.tr <- 1:length(s) %>%
-  sapply(function(j){
-  sp <- as.character(s[j])
-  temp <- list()
-  for (i in 1:n_sampling_replicas){
-    int <- sample(1:nLoc(pelo), s[j], replace = FALSE)
-    s_pelo <- pelo[, int]
-    if (i == 1)
-    boots_trees <- bootstrap_nj(s_pelo, nboot, type = "snp")
-    if (i > 1)
[…]

### 2019-05-09 code/6_Prepare_structure.R

@@ -1,68 +1,0 @@
-###.............................................................................
-# (c) Miguel Camacho Sánchez
-# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
-# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
-# April 2019
-###.............................................................................
-#GOAL: export data into structure format
-#DESCRIPTION: export data into structure format and create params files
-#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
-###.............................................................................
-#  REQUIRED FILES:
-#   Description:
-#   Inpath:
-#  OUTPUT:
-#    Description: raw genotypes plus metadata for dartseq samples
-#    Outpath:
-#  DEPENDENCIES:
-###.............................................................................
-library(dartR)
-library(dplyr)
-
-#1. Create STRUCTURE input files from dart and usat genotypes
-input_path <- paste0("data/intermediate/filt_genotypes.rds")
-gen <- readRDS(file = input_path)
-source("code/functions/edit_params.r")
-
-#warning dartR::gi2gl %>% gl2structure does not format format properly
-#structure files. But, hierfstat::write.struct has a bug so that it does not
-#print individual names in the ilab argument. I overcame this by modifiying the
-#dataframe with genotypes:
-names(gen) %>% seq_along() %>%
-  sapply(function(x){
-    dir_str <- file.path("data/intermediate", paste0("str_", names(gen)[x]))
-    dir.create(dir_str, showWarnings = T)
-    str_file <- file.path(dir_str, paste0(names(gen)[x], ".str"))
-    edit_mainparams(gl = gen[[x]], name = names(gen)[x],
-      str_file = str_file, dir_str = dir_str)
-    edit_extraparams(gl = gen[[x]], name = names(gen)[x],
-      str_file = str_file, dir_str = dir_str)
-    if (class(gen[[x]]) == "genind"){
-      hierfstat::genind2hierfstat(gen[[x]]) %>%
-      dplyr::mutate(pop = as.factor(indNames(gen[[x]]))) %>%
-      hierfstat::write.struct(fname = str_file)
-      } else if (class(gen[[x]]) == "genlight") {
-      dartR::gl2structure(gen[[x]], outfile = str_file)
-    }
-  })
-
[…]

### 2019-05-09 code/6_export_structure.R

@@ -1,54 +1,0 @@
-###.............................................................................
-# (c) Miguel Camacho Sánchez
-# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
-# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
-# April 2019
-###.............................................................................
-#GOAL: export data into structure format
-#DESCRIPTION: export data into structure format and create params files
-#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
-###.............................................................................
-#  REQUIRED FILES:
-#   Description:
-#   Inpath:
-#  OUTPUT:
-#    Description: raw genotypes plus metadata for dartseq samples
-#    Outpath:
-#  DEPENDENCIES:
-###.............................................................................
-library(dartR)
-library(dplyr)
-
-input_path <- paste0("data/intermediate/filt_genotypes.rds")
-gen <- readRDS(file = input_path)
-source("code/functions/edit_mainparams.r")
-
-#warning dartR::gi2gl %>% gl2structure does not format format properly
-#structure files. But, hierfstat::write.struct has a bug so that it does not
-#print individual names in the ilab argument. I overcame this by modifiying the
-#dataframe with genotypes:
-names(gen) %>% seq_along() %>%
-  sapply(function(x){
-    dir_str <- file.path("data/intermediate", paste0("str_", names(gen)[x]))
-    dir.create(dir_str, showWarnings = T)
-    file.copy("data/raw/extraparams", file.path(dir_str, "extraparams"),
-      overwrite = T) #same for all runs
-    file.copy("data/raw/mainparams", file.path(dir_str, "mainparams"),
-      overwrite = T) #edit for each run
-    str_file <- file.path(dir_str, paste0(names(gen)[x], ".str"))
-    edit_mainparams(gl = gen[[x]], name = names(gen)[x],
-      str_file = str_file, dir_str = dir_str)
-    if (class(gen[[x]]) == "genind"){
-      hierfstat::genind2hierfstat(gen[[x]]) %>%
-      dplyr::mutate(pop = as.factor(indNames(gen[[x]]))) %>%
-      hierfstat::write.struct(fname = str_file)
-      } else if (class(gen[[x]]) == "genlight") {
-      dartR::gl2structure(gen[[x]], outfile = str_file)
-    }
-  })
-
-
-#   cp extraparams to all subfolders
[…]

### 2019-05-09 code/functions/edit_mainparams.r

@@ -1,25 +1,0 @@
-#edit mainparams file for each STRUCTURE run
-edit_mainparams <- function(gl = gl, name = name,
-  str_file = str_file, dir_str = dir_str){
-  source("code/parameters/structure.r")
-  h <- c("dart_hyla", "dart_pelo", "usat_hyla", "usat_pelo")
-  assertthat::assert_that(name %in% h, msg = "check name in edit_mainparams")
-  param <- readLines("data/raw/mainparams")
-  param[16] <- sub("xx",  10, param[16])#number of individual
-  param[27] <- sub("xx",  adegenet::nInd(gl), param[27])#number of individual
-  param[28] <- sub("xx",  adegenet::nLoc(gl), param[28])#number of loci
-  param[22] <- sub("xx",  str_file, param[22])#infile
-    output <- file.path("data/final", paste0("str_", name, "_run"))
-  param[23] <- sub("xx",  output, param[23])#outfile
-  if (grepl("dart*", name)){
-    param[17] <- sub("xx", burnin_snp, param[17])#burnin length
-    param[18] <- sub("xx", run_snp, param[18])#run length
-  } else if (grepl("usat*", name)){
-    param[17] <- sub("xx", burnin_usat, param[17])#burnin length
-    param[18] <- sub("xx", run_usat, param[18])#run length
-  }
-  assertthat::assert_that(all(grepl("xx", param) == F),
-    msg = "wrong replacement in mainparams file")
-  mainparams_path <- file.path(dir_str, "mainparams")
-  writeLines(param, mainparams_path)
-}

### 2019-05-09 code/01_read_data.R
metadata fixed

@@ -35,14 +35,14 @@ gen_dart %<>% metadata_Dartseq_Hyla() %>% metadata_Dartseq_Pelobates()
 #2. microsatellite
 source("code/functions/read_usats.r")
   #genlight objects with metadata
-hyla_usat <- read_Hyla_molleri()
-pelo_usat <- read_Pelobates_cultripes()
+hyla_usat <- read_Hyla_molleri() #outputs metadata_hyla to global env
+pelo_usat <- read_Pelobates_cultripes()#outputs metadata_pelo to global env
 
 #create objects for exporting
 gen <- list(dart_hyla = gen_dart[[1]], dart_pelo = gen_dart[[2]],
    usat_hyla = hyla_usat, usat_pelo = pelo_usat)
-metadata_dartseq <- list(metadata_hyla = metadata_hyla,
-  metadata_pelo = metadata_hyla)
+metadata_dartseq <- list(metadata_darthyla = metadata_darthyla,
+  metadata_dartpelo = metadata_dartpelo)
 #note: usat metadata is already stored inside genind objects.
 
 saveRDS(gen, file = "data/intermediate/raw_genotypes.rds")

### 2019-05-09 code/02_genotype_filtering.R

@@ -17,6 +17,7 @@
 ###.............................................................................
 library(dartR)
 library(dplyr)
+library(magrittr)
 
 source("code/functions/gl_stats.r")
 source("code/functions/ind_miss.r")
@@ -63,6 +64,33 @@ names(gen_filt) <- genfiltnames
 filt_stats <- geno_stats(genlist = gen_filt, data = "filt")
 
 ######################
+#remove individuals from metadata that were filtered in genotypes
+  meta_path <- paste0("data/intermediate/metadata_dartseq.rds")
+  metadata_dartseq <- readRDS(file = meta_path)
+#assert_that all ids in genotypes are present in metadata
+    assertthat::assert_that(all(indNames(gen_filt$dart_hyla) %in%
+      metadata_dartseq$metadata_darthyla$sample_id))
+    assertthat::assert_that(all(indNames(gen_filt$dart_pelo) %in%
+      metadata_dartseq$metadata_dartpelo$sample_id))
+    assertthat::assert_that(
+    #assert_that all ids unique
+    list(indNames(gen_filt$dart_hyla), indNames(gen_filt$dart_pelo),
+            metadata_dartseq$metadata_dartpelo$sample_id,
+            metadata_dartseq$metadata_darthyla$sample_id) %>%
+      sapply(function(x) all(duplicated(x)) == F) %>% all()
+    )
+    #remove individuals from metadata
+metadata_dartseq$metadata_darthyla %<>%
+  filter(sample_id %in% indNames(gen_filt$dart_hyla))
+metadata_dartseq$metadata_dartpelo %<>%
+  filter(sample_id %in% indNames(gen_filt$dart_pelo))
+    #create list with metadata for all samples
+  metadata_dartseq[[3]] <- gen$usat_hyla@other
+  metadata_dartseq[[4]] <- gen$usat_pelo@other
+  names(metadata_dartseq)[c(3, 4)] <- c("usat_hyla", "usat_pelo")
+assertthat::assert_that(length(metadata_dartseq) == length(gen),
+  msg = "metadata_all has a different length than gen_filt")
+#####################
 #Raw data
 saveRDS(raw_stats, paste0("data/intermediate/raw_dartseq_stats.rds"))
 #Filtered data
@@ -73,3 +101,4 @@ saveRDS(raw_stats, paste0("data/intermediate/raw_dartseq_stats.rds"))
   } #add usat genotypes to gen_filt object
[…]

### 2019-05-09 code/functions/map_distribution.r

@@ -9,16 +9,21 @@ shp_pelo <- rgdal::readOGR(dsn = p_pelo, layer = "data_0") %>%
   sp::spTransform(mapr@crs)
 shp_hyla <- rgdal::readOGR(dsn = p_hyla, layer = "data_0") %>%
   sp::spTransform(mapr@crs)
+  #check I loaded the right layers
+assertthat::assert_that(as.character(
+  shp_hyla@data$BINOMIAL)[1] == "Hyla arborea")
+assertthat::assert_that(as.character(
+  shp_pelo@data$BINOMIAL)[1] == "Pelobates cultripes")
 
   #dartseq locations
-  p_dartmeta <- ""
-  dartmeta <- readRDS()
+  p_dartmeta <- "data/intermediate/metadata_dartseq.rds"
+  dartmeta <- readRDS(p_dartmeta)
   #microsatellite locations
-  p_gen <- ""
-  gen <- readRDS()
-
-assertthat::assert_that(as.character(shp_hyla@data$BINOMIAL)[1] == "Hyla arborea")
-assertthat::assert_that(as.character(shp_pelo@data$BINOMIAL)[1] == "Pelobates cultripes")
+  p_gen <- "data/intermediate/filt_genotypes.rds"
+  gen <- readRDS(p_gen)
+  m_usat_hyla <- gen$usat_hyla$other
+  m_usat_pelo <- gen$usat_pelo$other
+  rm(gen)
 
 #Plot distribution
 plot(mapr, breaks = c(0, 500, 1000, 1500, 2000, 3500),

### 2019-05-09 code/functions/read_metadata.r

@@ -1,4 +1,4 @@
-#read metadata hyla
+#read metadata for dartseq samples
 metadata_Dartseq_Hyla <- function(x){
   f <- "data/raw/metadata_hyla_dartseq.xlsx"
   m <- readxl::read_excel(f)
@@ -16,7 +16,20 @@ assertthat::assert_that(!is.null(c(m[["dartseqID"]],
   pop(x$hyla) <- sapply(g_dartseq_id, function(x){
     which(m[["dartseqID"]] == x) %>% m[["locality"]][.]
   })
-  metadata_hyla <<- m
+  metadata_darthyla <<- m %>%
+              select(id, locality, latitude, longitude) %>%
+              mutate(latitude = as.numeric(latitude)) %>%
+              mutate(longitude = as.numeric(longitude)) %>%
+              filter(id %in% indNames(x$hyla)) %>%
+              rename(sample_id = id) %>%
+              #convert to spatial object
+              {sp::SpatialPointsDataFrame(data = .[, c(1, 2),
+              drop = FALSE], coords = .[, c("latitude", "longitude")])} %>%
+              as("sf") %>%
+              sf::st_set_crs(4326)
+  assertthat::assert_that(identical(
+    x = sort(as.character(indNames(x$hyla))),
+    y = sort(as.character(metadata_darthyla$sample_id))))
   return(x)
 }
 
@@ -27,7 +40,11 @@ metadata_Dartseq_Pelobates <- function(x){
     rename(full_locality = site) %>%
     rename(dartseqID = vaucher) %>%
     rename(latitude = lat) %>%
-    rename(longitude = lon)
+    rename(longitude = lon) %>% dplyr::as_tibble() %>%
+    {sp::SpatialPointsDataFrame(data = .[, c(1, 6), drop = FALSE],
+                                coords = .[, c("latitude", "longitude")])} %>%
+    as("sf") %>%
+    sf::st_set_crs(4326)
 
   assertthat::assert_that(class(x$pelo) == "genlight")
   m_darseq_id <- m[["dartseqID"]]
@@ -39,6 +56,8 @@ metadata_Dartseq_Pelobates <- function(x){
   pop(x$pelo) <- sapply(g_dartseq_id, function(x){
     which(m[["dartseqID"]] == x) %>% m[["locality"]][.]
   }) %>% unlist()
-  metadata_pelo <<- m
+  metadata_dartpelo <<- m %>% #select only samples present in the genlight object
[…]

### 2019-05-09 code/functions/read_usats.r

@@ -38,6 +38,10 @@ read_Pelobates_cultripes <- function() {
                      stringsAsFactors = FALSE)
         })
     }) %>%
+    mutate(Locality = as.character(Locality) %>% #simplify pop names
+    gsub("^[0-9].*a, ", "", .) %>%
+    gsub(",.*$", "", .) %>%
+    gsub("‐.*$", "", .))
     tibble::as_tibble()
   ## add in population data,
   ### since we only have sample ids for individuals with mtDNA data,
@@ -46,9 +50,9 @@ read_Pelobates_cultripes <- function() {
   ### as sample B, and we known that sample B is in population A then we know
   ### that sample A is in population A
   s2 <- tibble::tibble(sample_id = rownames(o@tab),
-                       population_id = as.character(o@pop)) %>%
+                       locality = as.character(o@pop)) %>%
        dplyr::left_join(s, by = c("sample_id" = "Sample.Code")) %>%
-       plyr::ddply("population_id", function(x) {
+       plyr::ddply("locality", function(x) {
          if (!all(is.na(x$ID))) {
            x$ID <- na.omit(x$ID)[[1]]
            x$Locality <- na.omit(x$Locality)[[1]]
@@ -62,7 +66,7 @@ read_Pelobates_cultripes <- function() {
                          by = "sample_id")
   ### fortunately it appears that only one population did not have any mtDNA
   ### samples so we can manually define the  popuation ids for this population
-  assertthat::assert_that(dplyr::n_distinct(s2$population_id[is.na(s2$ID)]) ==
+  assertthat::assert_that(dplyr::n_distinct(s2$locality[is.na(s2$ID)]) ==
                           1)
   missing_population <- setdiff(unique(s$ID), unique(s2$ID[!is.na(s2$ID)]))
   assertthat::assert_that(assertthat::is.number(missing_population))
@@ -81,7 +85,7 @@ read_Pelobates_cultripes <- function() {
   o@other <-
     s2 %>%
     dplyr::select(sample_id, Locality, Longitude, Latitude) %>%
-    dplyr::rename(population_id = Locality, x = Longitude, y = Latitude) %>%
+    dplyr::rename(locality = Locality, x = Longitude, y = Latitude) %>%
     {sp::SpatialPointsDataFrame(data = .[, c(1, 2), drop = FALSE],
[…]

### 2019-05-10 code/functions/centroid_sf.r
map done

@@ -1,0 +1,30 @@
+#calculate centroid from a sf object given an index (default = "locality")
+centroid_sf <- function(sf_object = sf_object, index = "locality",
+                        latlon = c("longitude", "latitude")){
+  sfc_as_cols(sf_object) %>%
+  plyr::ddply(index, function(x){
+  x <- dplyr::select(x, longitude, latitude)
+  x <- sp::SpatialPoints(x, proj4string = CRS("+init=epsg:4326"))
+  x <- rgeos::gCentroid(x)
+  x <- as.data.frame(x)
+  names(x) <- latlon
+  x
+  }) %>%
+  {sp::SpatialPointsDataFrame(data = .[, 1, drop = FALSE],
+        coords = .[, latlon])} %>%
+  as("sf") %>%
+  sf::st_set_crs(4326)
+}
+
+
+sfc_as_cols <- function(x, names = c("longitude", "latitude")) {
+  stopifnot(inherits(x, "sf") && inherits(sf::st_geometry(x), "sfc_POINT"))
+  ret <- sf::st_coordinates(x)
+  ret <- tibble::as_tibble(ret)
+  stopifnot(length(names) == ncol(ret))
+  x <- x[, !names(x) %in% names]
+  ret <- setNames(ret, names)
+  x <- dplyr::bind_cols(x, ret)
+  st_geometry(x) <- NULL
+  x
+}

### 2019-05-10 code/functions/map_distribution.r

@@ -1,7 +1,11 @@
 #create map with the distribution of the samples and native distribution of the species
+library(dplyr)
+library(sf)
+library(raster)
 #read raster and shape files
-p_raster <- "data/raw/raster_iberian_peninsula.rds"
-mapr <- readRDS(p_raster)
+  #raster
+p_raster <- "data/raw/raster_iberian_peninsula.grd"
+mapr <- raster::raster(p_raster)
   #IUCN distribution
 p_pelo <- "data/raw/redlist_species_data_44f009a9-b3b2-4608-a818-24c3dd714900"
 p_hyla <- "data/raw/redlist_species_data_eff4cf11-63bb-4131-bba3-104c3358697b"
@@ -9,37 +13,46 @@ shp_pelo <- rgdal::readOGR(dsn = p_pelo, layer = "data_0") %>%
   sp::spTransform(mapr@crs)
 shp_hyla <- rgdal::readOGR(dsn = p_hyla, layer = "data_0") %>%
   sp::spTransform(mapr@crs)
-  #check I loaded the right layers
+    #check I loaded the right layers
 assertthat::assert_that(as.character(
   shp_hyla@data$BINOMIAL)[1] == "Hyla arborea")
 assertthat::assert_that(as.character(
   shp_pelo@data$BINOMIAL)[1] == "Pelobates cultripes")
 
-  #dartseq locations
-  p_dartmeta <- "data/intermediate/metadata_dartseq.rds"
-  dartmeta <- readRDS(p_dartmeta)
-  #microsatellite locations
-  p_gen <- "data/intermediate/filt_genotypes.rds"
-  gen <- readRDS(p_gen)
-  m_usat_hyla <- gen$usat_hyla$other
-  m_usat_pelo <- gen$usat_pelo$other
-  rm(gen)
+  #sample locations
+  p_meta <- "data/intermediate/metadata_all.rds"
+  meta <- readRDS(p_meta)
+#load functions
+source("code/functions/centroid_sf.r")
 
 #Plot distribution
-plot(mapr, breaks = c(0, 500, 1000, 1500, 2000, 3500),
-      col = gray.colors(5, start = 0.9, end = 0.3)) #issue margins
-plot(shp_pelo, add = T, border = "black",
-  col = grey(0.5, alpha = 0.3))
-title(shp_pelo@data$BINOMIAL[1])
+outfile <- paste0("data/final/map_samples.pdf")
[…]

### 2019-05-10 code/functions/read_metadata.r

@@ -4,12 +4,12 @@ metadata_Dartseq_Hyla <- function(x){
   m <- readxl::read_excel(f)
   m_darseq_id <- m[["dartseqID"]]
   assertthat::assert_that(class(x$hyla) == "genlight")
-  g_dartseq_id <- indNames(x$hyla) %>% gsub(" ", "", .)
+  g_dartseq_id <- adegenet::indNames(x$hyla) %>% gsub(" ", "", .)
 assertthat::assert_that(!is.null(c(m[["dartseqID"]],
       m[["locality"]],
       m[["id"]])))
   #fix sample names
-  indNames(x$hyla) <- sapply(g_dartseq_id, function(x){
+  adegenet::indNames(x$hyla) <- sapply(g_dartseq_id, function(x){
     which(m[["dartseqID"]] == x) %>% m[["id"]][.]
   })
   #fix pop names
@@ -20,15 +20,15 @@ assertthat::assert_that(!is.null(c(m[["dartseqID"]],
               select(id, locality, latitude, longitude) %>%
               mutate(latitude = as.numeric(latitude)) %>%
               mutate(longitude = as.numeric(longitude)) %>%
-              filter(id %in% indNames(x$hyla)) %>%
+              filter(id %in% adegenet::indNames(x$hyla)) %>%
               rename(sample_id = id) %>%
               #convert to spatial object
               {sp::SpatialPointsDataFrame(data = .[, c(1, 2),
-              drop = FALSE], coords = .[, c("latitude", "longitude")])} %>%
+              drop = FALSE], coords = .[, c("longitude", "latitude")])} %>%
               as("sf") %>%
               sf::st_set_crs(4326)
   assertthat::assert_that(identical(
-    x = sort(as.character(indNames(x$hyla))),
+    x = sort(as.character(adegenet::indNames(x$hyla))),
     y = sort(as.character(metadata_darthyla$sample_id))))
   return(x)
 }
@@ -42,13 +42,13 @@ metadata_Dartseq_Pelobates <- function(x){
     rename(latitude = lat) %>%
     rename(longitude = lon) %>% dplyr::as_tibble() %>%
     {sp::SpatialPointsDataFrame(data = .[, c(1, 6), drop = FALSE],
-                                coords = .[, c("latitude", "longitude")])} %>%
+                                coords = .[, c("longitude", "latitude")])} %>%
     as("sf") %>%
     sf::st_set_crs(4326)
 
[…]

### 2019-06-18 code/rmarkdown/article.Rmd
k finder fix and introduction

@@
 **general overview about the usat vs snp**
-1-which are the characteristics of each marker
-2-how informative are they
-3-pros and cons for both
-4-limitations
-5-previous studies focusing on usats vs SNPs
-**our case study**
-1-study species: characteristics, distrigution, why them
-2-markers datasets
-**brief methodology with results and conclusions**
+Genetic markers represent a central pillar upon which Conservation and Evolutionary Biology are built.
+Ideally, neutral genetic markers should be representative of genome-wide processes and have  adequate levels of information to address the desired demographic or evolutionary questions.
+Under these premises, the popularity of different types of neutral genetic markers have been oscillating since the advent of molecular biology.
+While short-sequence tandem repeats (microtellites hereafter) have been at the frontline of several biological disciplines since the past xx years, the rise of high-throughput sequencing in the last decade has pushed single nucleotide polymorphisms (SNPs) at this forefrontline as well, and, at present, these are two of the most widely used markers in xx research.
+Each have different characteristics deriving from their molecular nature.
+Microsatellites are repeats of short-motifs (xx -- bp) in the genome caused by slippage of the DNA-polymerase, whereas SNPs are the form of single nucleotide (ATCG).
+Due to the relatively high rate of polymerase slippage compared to base misincorporation, microtellites have mutation rates which are several orders of magnitude higher than SNPs.
+For instance, in vertebrates??
+the mutation rate can vary from xx to xxx, whereas the background mutation rate for vertebrates genomes (often used for mutation rate in SNPs, assuming they are randomly distributed across the genome) is around 10-9.
+The number of alleles for a given microsatellite can be very high due to the presumably infinite number of repeats (assuming an infinite allele site model) it could adopt.
+However, SNPs are restricted to four different alleles corresponding to the 4 nucleotides (ATCG), and often only biallelic SNPs are considered in most studies.
+Microtellites have been a preferable marker choice for population genetics since xxx years, since they claim to be more sensitive to demographic processes at shallow evolutionary scales due to their high levels of polymorphism and high mutation rates.
+On the contrary, SNPs contain little information less information per locus.
+The rise of high-throughput sequencing in the last decade has pushed SNPs to the frontline in population genetics.
+The lesser per-locus information in SNPs is compensated by the sequencing of thousands of them at the same cost than for microsatellites.
+Approximately, developeing and sequencing a panel of xxx microsatellites has the same cost than xxx SNPs (Hodel 2016??).
+Genome-wide SNPs secure a distribution of mutation rates, from low to high, which can provide information and different evolutionary scales from recent within species processes to interspecies, contrary to microsatellites for which homoplasies and the difficulties to fit an evolutionary model only make them suitable for whithin species-processes.
+
+Whereas the value of microtellites has been largely proven for population genetics, the attemps to scale their information into a phylogenetics framework (ie phylogeography) have had little success.
+Indeed, the literature is full of attempts to recover within species phylogenetic reconstructures which recurrently fall into the issue of being supported.
+Recent studies have tried to make comparisons on the levels of iformation returned for microsatelllites versus SNPs for the sampe study system.
+They find......
+However, when referening to comparisons for the genetic structure, they often refer terms "resolution" based on visual examination of the results between different marker, but they do not provide any statistic.
+Also, they have not dealt with the issue of the information that they can provide in a phylogenetics framework.
+We evaluated the genetic structure and phylogeography information yielded by comparable datasets of microtellites and SNPs for two amphibians from the Iberian Peninsula, *Pelobates cultripes* and *Hyla molleri*.
+??
+aims???
+hypotheses??
+We used microtellite datasets from published studies, Gutiérrez-Rodríguez et al.
+(2017) (for *P.
+cultripes*) and Sánchez-Montes et al.
+(2019), for *H.
+molleri*, and newly sequenced panels of SNPs.
+Our tree reconstructions underline the strong limitations of using microsatellites for phylogenetics, specially for getting information at deeper intra-specific processes, contrary to SNPs, which proved useful across all depths in the trees.
+For population structure xxxxx.
 

### 2019-06-18 code/03_structure/08_kfinder.R

@@ -15,23 +15,36 @@
 #    Outpath:
 #  DEPENDENCIES:
 ###.............................................................................
-
+library(dplyr)
+#read file with parameters
 #1. create parameters file
 # vector with folder names for each run
 runs <- dir("data/final", pattern = "run_", full.names = T)
 
 runs %>%
   sapply(function(x){
+    #get k values
+    ks <-
+      dir(x) %>%
+        stringr::str_extract(pattern = "K[0-9]+") %>%
+          stringr::str_extract(pattern = "[0-9]+") %>%
+          as.numeric %>% range(na.rm = T)
+    #get number of reps
+    reps <-
+      dir(x) %>%
+        stringr::str_extract(pattern = "rep[0-9]+") %>%
+          stringr::str_extract(pattern = "[0-9]+") %>%
+          as.numeric %>% max(na.rm = T)
     #create parameter file
     param <- c(sprintf('"%s"', file.path(getwd(), x)),
                sprintf('"%s"', "str*"),
-               1, #from K
-               8, #to K
+               ks[1], #from K
+               ks[2], #to K
                readLines(file.path(x, "K2_rep1.stlog")) %>%
                  grep(pattern = " individuals$", value = T) %>%
                  stringr::str_extract("[0-9]+") %>%
                  as.numeric(),
-               5) #number of replicates per K
+               reps) #number of replicates per K
     fp1 <- file.path(x, "kfinder.par") #file path of parameter file
     writeLines(con = fp1, text = param) #write parameter file
 #2. run kfinder within each run folder

### 2019-06-18 code/render.R

@@ -1,0 +1,3 @@
+#render article
+rmarkdown::render("code/rmarkdown/article.Rmd",
+  output_file = "usat_snp.docx", output_dir = "article")

### 2019-06-26 code/rmarkdown/Figures.rmd
intro and Fig1

@@
 ##Figure 1:
-```{r, fig.height = 4.5, fig.width = 4.5, fig.cap = "Distribution and populations sampled in *Hyla molleri* (panel A) and *Pelobates cultripes* (panel B), with corresponding genotype dataset: microsatellite (triangles)  and SNP (circles).
-The number inside each symbol refer to the population ID for microsatellite data in Sánchez-Montes et al.
-(2019)(*H.
-molleri), Gutiérrez-Rodríguez et al.
-(2017)(*P.
-cultripes) and SNPs Table 1.
-The sample sizes for each population are indicated in parenthesis.
-The distribution for each species is indicated in shaded area, downloaded from the IUCN on May 2019.(xx note: because H.
-molleri is not in the IUCN database still, I used the distribution of H.
-arborea)"}
+```{r, fig.height = 4.5, fig.width = 4.5, fig.cap = "Populations sampled in *Hyla molleri* (panel A) and *Pelobates cultripes* (panel B), for microsatellite (triangles) and SNP (circles).
+Coordinates and samples codes are in Table 1.
+The distribution for each species is indicated in shaded area, downloaded from the IUCN on May 2019 (*H.
+molleri* was described recently and its distribution is not availabe at the IUCN database.
+We used the distribution of *H.
+arborea* instead)"}
 

### 2019-06-26 code/rmarkdown/article.Rmd
intro and Fig1

--- 2019-05-09 code/rmarkdown/article.Rmd
@@
 # 1. Introduction
-**general overview about the usat vs snp**
-1-which are the characteristics of each marker
-2-how informative are they
-3-pros and cons for both
-4-limitations
-5-previous studies focusing on usats vs SNPs
-**our case study**
-1-study species: characteristics, distrigution, why them
-2-markers datasets
-**brief methodology with results and conclusions**
+
+Microsatellites furnished at the 90's a powerful tool in ecology and conservation biology to assess variation at nuclear loci for the study of biological processes at the population level (Jehleh and Artzen 2002; Allendorf 2016).
+Their popularity has remained constant until present, despite the irruption and continuous growth in the use of SNPs since the 2000's (Putman and Carbone 2014; Allendorf 2016).
+Mutation rates in microsatellites are several orders of magnitude higher compared to SNPs across taxa (Dallas 1992; Weber and Wong 1993; Zhang & Hewitt, 2003; Lynch 2010).
+The high mutation rates combined with the large number of different a alleles for a single locus generate immense levels of polymorphism useful in population genetics (Avise 2004; Allendorf 2016).
+An immediate consequence is that microtellites are claimed to be more sentitive to sudden demographical processes (Morin 2004), so they could have more power to detect subtle population structure or recent bottlenecks (Luikart et al.
+1998; Putman and Carbone 2014; Haasl and Payseur 2010).
+However, these premises have been questioned (Peery et al.
+2012).
+High polymorphism comes together with high homoplasy (Garza and Freimer 1996; Hedrick 1999) and the difficulties to fit complex evolutionary models (Valdes et al.
+1993; Rienzo et al 1994).
+The precedent can lead to estimate unreliable genetic distances for large divergence times (Kalinowski 2002), underestimate differentiation between populations caused with high intra-populational heterozigosity or overestimation of divergence caused by recent bottlenecks (Hedrick 1999).
+The simple evolutionary models in SNPs, high abundance and widespread distribution across the genome promised to overcome the limitations of microtellites in population genetics (Morin 2004).
+But, a microsatellite locus contains from 4 to 12 times more information than a SNPs (Liu et al.
+2005), and the high initial costs of SNP discovery and sequencing (Morin 2004) constrained the use of SNPs to tens in non-model species or to thousands only in model organisms and humans (Liu et al.
+2005; Pucket 2017).
+The first studies comparing few microsatellites to tens of SNPs found that the performance between both markers varied from study system and questions addressed (Puckett 2017), but that often, the few informative SNPs yielded the same amount of information than the microsatellite dataset (Liu et al.
+2005; Narum et al.
+2008; Hess et al.
+2011).
+The development of Genotyping-by-Sequencing techniques, such as RADseq (Miller et al.
+2007; Baird et al.
+2008) and other similar techniques of reduction of genome complexity, coupled with the cost reduction of massive parallel sequencing has extended the reach of massive SNP genotyping to non-model organisms (Baird et al.
+2008; Davey et al.
+2011; Peterson et al.
+2012).
+This has boosted the studies with thousands of SNPs in wildlife organism in conservation and evolutionary biology (Putman and Carbone 2014; Allendorf 2016; Andrews et al.
+2016).
+The lesser per-locus information in SNPs is largely compensanted by the sequencing of thousands of them at the same cost than for a handful of microsatellites (Hodel et al.
+2017; Puckett 2017).
+In the current genomic era, studies on wildlife organisms using thousands of SNPs typically find they being more powerful than microsatellites (Puckett 2017).
+So, yielding more defined clustering in non-parametric clustering (eg.
+PCA or DAPC; Rasic et al.
+2014; Candy et al.
+2015; Malenfant et al.
+2015; Elbers et al.
+2016; Jeffries et al.
+2016; Puckett and Eggert 2016; Hodel et al.
+2017), higher resolution to detect hybridization (Bradbury et al.
+2015), better explanation for patterns of isolation by distance (Jeffries et al.
+2016), more accuracy for estimating relatedness or origin of individuals (Puckett and Eggert 2016; Thrasner et al.
+2018; Lemopoulos et al.
+2019) or better trees (Bruneaux et al.
+2012; Candy et al.
+2015).
+The above-mentioned sharp clustering with large-scale SNP data compared to microtellites using non-parametric methods does not reflect when using model-based clustering in the popular software STRUCTURE.
+This software provides a powerful tool to estimate individual ancestries from genotypes data (Pritchard et al.
+2000).
+Simulations show equivalent power of tens of microtellites and thousands of SNPs to retrieved with high confidence the correct number of cluster (Haasl and Payseur 2011).
+With empirical data, overall genetic structure with both datasets can vary in the number of most optimal number of clusters (best K) and proportions of individual ancestries (Bradbury et al.
+2015; Malenfant et al.
+2015; Elbers et al.
+2016; Bohling et al.
+2019) or be practically identical (Lemopoulos et al.
+2019).
+However, the best K can be very sensitive to demographic processes and can vary substantially depending on the method to estimate it (Janes et al.
+2017; Wang 2019).
+Other issues as uneven sampling and hierarchical population structure can also difficult the inference of the correct population structure (Puechmaille et al.
+2016; Wang 2016).
+Neighbor joining clustering is a complementary way to overcome some limitations of model-based clustering and to recover a reliable evolutionary history of the populations under a wide variaty of scenarios (Kalinowski 2009, 2011).
+The reconstructions of phylogenetic trees using neighbor joining with microtellite data are often coherent with the phylogeographic structure recovered with SNPs.
+However, the large homoplasy undermines their power to retrieve phylogenetic relationships between distant populations (Estoup et al.
+2002).
+So, trees from microsatellite data recurrently have little support, specially at deeper nodes, which limits the interpretations of the evolutionary history of the populations (Moore et al.
+2014; Candy et al.
+2015; Bruneaux et al.
+2012; Bradbury et al.
+2018), although when several hundreds of loci are used the correct topology is recovered correctly (Takezaki and Nei 2006).
+The large amount of SNPs and their genome-wide distribution secure a range mutation rates which can provide sufficient information at different evolutionary scales, from recent within species processes to interspecies phylogenies (eg.
+Defaveri et al.
+2013; Petersen et al.
+2013).
+
+We aimed to evaluate the phylogeographic patterns yielded by comparable datasets of microtellites and SNPs for two amphibians from the Iberian Peninsula, *Pelobates cultripes* and *Hyla molleri*.
+We hypothesize that the recent evolutionary window of microsatellites compared to the larger window in SNPs should lead to (1) model-based clustering for microsatellites yielding a higher number of clusters and less individual admixture and (2) very limited phylogeographic resolution of microtellite data to shallow evolutionary scales compared to a more robust phylogeographic resolution of SNPs at all scales.
+We used microtellite datasets from published studies, Gutiérrez-Rodríguez et al.
+(2017) (for *P.
+cultripes*) and Sánchez-Montes et al.
+(2019), for *H.
+molleri*, and newly sequenced panels of SNPs using DArTseq.
+Our tree reconstructions underline the strong limitations of typical microsatellite panels for phylogenetics, specially for getting information at deeper intra-specific processes, contrary to SNPs, which proved useful across all intra-species evolutionary scales.
 
@@
 ## 2.1 Data collection
-Our analysis are built upon microsatellite and SNP datasets for two Iberian amphibians: *Hyla molleri* and *Pelobates cultripes*.
-For each species, the microtellite and SNP datasets come from different sample sets but covering both most of their native range (Figure 1).
-The microsatellite genotypes were taken from already-published studies.
+We used microsatellite and SNP datasets from two Iberian amphibians, *Hyla molleri* and *Pelobates cultripes*, covering most of their native range (Figure 1).
+The microsatellite genotypes were obtained from already-published studies.
 For *Hyla molleri*, we gathered the microsatellite genotypes from *Appendix 1* in Sánchez-Montes et al.
@@
 This dataset consisted of `r adegenet::nInd(gen$usat_pelo)` samples with `r adegenet::nLoc(gen$usat_pelo)` genotyped microsatellites.
-The SNP data was generated using DArTseq technology (xx).
+The SNP data was generated using DArTseq technology (Diversity Arrays Technology, Canberra, Australia).
+We included samples spanning all over the Iberian Peninsula (Figure 1; Table 1).
+DNA was extracted ExtractMe Genomic DNA 96-Well (DNA GDAŃSK), concentrated with QIAamp DNA Micro (QIAGEN GmbtH) kits, if necessary, and sent for sequencing.
 ##

### 2019-06-26 code/01_prepare_data/04_Filter_populations.R

@@ -137,7 +137,5 @@ gen %<>% lapply(function(x){
   x
   })
 
-#Plot samples post filtered pop
-plot_samples(meta = gen,
-  name_plot = "data/intermediate/sample_postfilt_distribution.pdf")
+#Save object
 saveRDS(gen, "data/intermediate/gen_consolidated_filtered.rds")

### 2019-06-26 code/Figure1_map_samples.R

@@ -1,0 +1,16 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# June 2019
+###.............................................................................
+#GOAL: Plot map wih samples
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+#plotting function
+source("code/functions/map_distribution.r")
+#load consolidated genotypes
+gen <- readRDS("data/intermediate/gen_consolidated_filtered.rds")
+
+plot_samples(meta = gen,
+  name_plot = "data/final/Figure1_map_samples.pdf", res = 50000)

### 2019-06-26 code/functions/map_distribution.r

@@ -1,8 +1,11 @@
-plot_samples <- function(meta = metadata, name_plot){
+plot_samples <- function(meta = metadata, name_plot, res = res,
+  mode = c(1, 2)){
 #meta is a metadata list where the name of each element has names of marker
 #and species. Each element is a sf object with #sample_id#locality#geometry
 #create map with the distribution of the samples and native
 #distribution of the species
+#res, resolution of raster. For fast plotting use 5.10^3. For high-res 5.10^4
+#mode is 1 for test mode and 2 for pretty plotting
 library(dplyr)
 library(sf)
 library(raster)
@@ -50,10 +53,10 @@ for (i in seq_along(x)){
   centroids_dart <- centroid_sf(
     meta[[grep(paste0("dart.", x[i]), names(gen))]]@other$metadata)
   #plot
-  plot(mapr, maxpixels = 5000, #raster
+  plot(mapr, maxpixels = res, #raster
        breaks = c(0, 500, 1000, 1500, 2000, 3500),
        col = gray.colors(5, start = 0.9, end = 0.3),
-       axes = T, box = FALSE)
+       axes = T, box = FALSE, ylab = "Latitude", xlab = "Longitude")
   plot(get(ls()[grepl(paste0("shp_", x[i]), ls())]),  #IUCN distribution
        add = T, border = "black", col = grey(0.5, alpha = 0.3))
   plot(sf::st_geometry(centroids_dart), add = T, pch = pch_snp, cex = 1.4,
@@ -62,8 +65,9 @@ for (i in seq_along(x)){
        col = micro_col) #usats samples
   title(y[i], font.main = 3, cex.main = 0.8)
   if (i == 1){
-    legend("topleft", legend = c("Microsatellites", "SNPs"),
-           pch = c(pch_usat, pch_snp), bty = "n", col = c(micro_col, snp_col))
+    legend(y = extent(mapr)@ymax, x = extent(mapr)@xmin,
+      legend = c("Microsatellites", "SNPs"), cex = 0.7,
+      pch = c(pch_usat, pch_snp), bty = "n", col = c(micro_col, snp_col))
   }
   off <- 0.5
   pop_usat <- centroids_usat %>% sfc_as_cols %>%
@@ -72,10 +76,16 @@ for (i in seq_along(x)){
   pop_dart <- centroids_dart %>% sfc_as_cols %>%
     mutate(longitude = as.numeric(longitude)) %>%
[…]

### 2019-06-28 code/rmarkdown/article.Rmd
table 1

@@
 output:
-  rmarkdown::pdf_document:
-    toc: false
-    keep_tex: true
-    fig_caption: yes
-    includes:
-      in_header: preamble.tex
-fontsize: 11pt
-documentclass: article
-subparagraph: true
+  word_document: default
+  pdf_document: default
 ---

### 2019-06-28 code/rmarkdown/tables.Rmd
table 1

---
title: "Tables"
output:
  word_document: default
  html_document:
    df_print: paged
---
Table 1. Samples and their localities. In the sample column, we indicate the species(*P.c.* and *H.m*, to refer to *Pelobates cultripes* and *Hyla molleri*, respectively) and marker type (SNPs or microsatellite).
```{r, eval = T, echo = F, include = T, warning = F, message=FALSE}
library(flextable)
library(dplyr)
h <- readRDS("../../data/intermediate/table1.rds")

ft <-
h %>% flextable::flextable() %>%
  flextable::width(j = 5, 3.7) %>%
  flextable::width(j = 2, 2) %>%
  flextable::fontsize(size = 8, part = "body") %>%
  flextable::fontsize(size = 10, part = "header")
ft
```

### 2019-06-28 code/Table1_samples.R

@@ -1,0 +1,97 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# June 2019
+###.............................................................................
+#GOAL: Table with samples
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+#load consolidated genotypes
+gen <- readRDS("data/intermediate/gen_consolidated_filtered.rds")
+
+library(dplyr)
+source("code/functions/centroid_sf.r")
+
+#one table with all data
+h <-
+seq_along(gen) %>%
+  lapply(function(x){
+    gen[[x]]@other$metadata %>%
+      sfc_as_cols %>% #sf object to dataframe with lat lon columns
+      dplyr::mutate(dataset = names(gen)[x]) %>% #add a column with dataset
+      #round coordinates
+      dplyr::mutate(longitude = round(longitude, 3)) %>%
+      dplyr::mutate(latitude = round(latitude, 3))
+  }) %>%
+  plyr::ldply(rbind) %>% #bind all tables within the list
+  tibble::as_tibble()
+
+#inspect duplicated samples
+dd <- h$sample_id[which(duplicated(h$sample_id))] #dublicated ids
+#arrange by sample id
+h %>% dplyr::filter(sample_id %in% dd) %>% dplyr::arrange(sample_id)
+
+#there are some incongruences in the coordinates (difference in a hundreth of a degree) for some of the duplicated samples. I have checked at the coordinates on the original metadata and the publications and everything seems to be fine, so I do not where the error comes from.
+
+#table 1: expanded including duplicates
+p <-
+  h %>%
+  dplyr::mutate(species = ifelse(grepl("hyla", dataset), "H.m.",
+                                ifelse(grepl("pelo", dataset), "P.c.", NA))) %>%
+  dplyr::mutate(marker = ifelse(grepl("dart", dataset), "SNPs",
+                     ifelse(grepl("usat", dataset), "microsatellites", NA))) %>%
+  select(-dataset) %>%
+  arrange(locality, sample_id)
+
+#table 2: compact
+
[…]

### 2019-06-28 code/functions/centroid_sf.r

@@ -4,7 +4,7 @@ centroid_sf <- function(sf_object = sf_object, index = "locality",
   sfc_as_cols(sf_object) %>%
   plyr::ddply(index, function(x){
   x <- dplyr::select(x, longitude, latitude)
-  x <- sp::SpatialPoints(x, proj4string = CRS("+init=epsg:4326"))
+  x <- sp::SpatialPoints(x, proj4string = sp::CRS("+init=epsg:4326"))
   x <- rgeos::gCentroid(x)
   x <- as.data.frame(x)
   names(x) <- latlon
@@ -16,7 +16,7 @@ centroid_sf <- function(sf_object = sf_object, index = "locality",
   sf::st_set_crs(4326)
 }
 
-
+#converts sf object to dataframe with lat lon columns
 sfc_as_cols <- function(x, names = c("longitude", "latitude")) {
   stopifnot(inherits(x, "sf") && inherits(sf::st_geometry(x), "sfc_POINT"))
   ret <- sf::st_coordinates(x)
@@ -25,6 +25,6 @@ sfc_as_cols <- function(x, names = c("longitude", "latitude")) {
   x <- x[, !names(x) %in% names]
   ret <- setNames(ret, names)
   x <- dplyr::bind_cols(x, ret)
-  st_geometry(x) <- NULL
+  sf::st_geometry(x) <- NULL
   x
 }

### 2019-06-28 code/render.R

@@ -1,3 +1,6 @@
 #render article
 rmarkdown::render("code/rmarkdown/article.Rmd",
   output_file = "usat_snp.docx", output_dir = "article")
+
+rmarkdown::render("code/rmarkdown/tables.Rmd",
+  output_file = "Tables.docx", output_dir = "article")

### 2019-07-01 code/rmarkdown/Supporting_information.rmd
convergence plots

@@
 ---
+Table S1: samples
+Figure S1: SNP filtering for *H.
+molleri* (A, C, E, G) and *P.
+cultripes* (B, D, F, H).
+En each panel we show the frequency distribution (y-axes) of the metric evaluated (x-axes), with the thresholds applied (dashed vertical lines).
+
+Figures 2.1 - 2.11: Change of the likelihood of the data across iterations for 10 replicate runs for K = 1 to K = 10. The vertical dashed lines indicate burnin length.
+
+Figures S3.1 - S3.2: Gelman and Rubin's convergence diagnostic for SNP (S3.1) and microtellite (S3.2) datasets.

### 2019-07-01 code/rmarkdown/article.Rmd
convergence plots

@@
 knitr::opts_knit$set(root.dir = normalizePath("../.."))
+library(dplyr)
 ```
@@
 ```{r, include = FALSE}
-# restore session
-gen <- readRDS("data/intermediate/raw_genotypes.RDS")
-# load packages
-# define helper functions
-# load data
+# data
+gen <- readRDS("data/intermediate/gen_consolidated_filtered.rds")
+samples <- readRDS("data/intermediate/table1.rds")
+
+#Results
+#number of individuals H. molleri
+indH <- names(gen) %>% grep(pattern = "hyla") %>%
+            sapply(function(x) adegenet::nInd(gen[[x]])) %>% sum
+#number of individuals P. cultripes
+indP <- names(gen) %>% grep(pattern = "pelo") %>%
+            sapply(function(x) adegenet::nInd(gen[[x]])) %>% sum
+indHm <- adegenet::nInd(gen$usat_hyla)
+indHs <- adegenet::nInd(gen$dart_hyla)
+indPm <- adegenet::nInd(gen$usat_pelo)
+indPs <- adegenet::nInd(gen$dart_pelo)
+#total individuals
+totalind <- indP + indH
+#localities for H. molleri
+popH <- samples$samples %>% grep(pattern = "H.m.") %>% length
+#localities for P. cultripes
+popP <- samples$samples %>% grep(pattern = "P.c.") %>% length
+#missing data microtellite
+# H. molleri
+missHm <- hierfstat::genind2hierfstat(gen$usat_hyla) %>%
+            dplyr::select(-1) %>% {xx <<- length(as.matrix(.)); . } %>%
+            is.na() %>% {sum(.) / xx * 100} %>% round(1)
+# P. cultripes
+missPm <- hierfstat::genind2hierfstat(gen$usat_pelo) %>%
+            dplyr::select(-1) %>% {xx <<- length(as.matrix(.)); . } %>%
+            is.na() %>% {sum(.) / xx * 100} %>% round(1)
+#missing data SNPs
+missHs <- as.matrix(gen$dart_hyla) %>%
+            {xx <<- length(.); . } %>%
+            is.na() %>% {sum(.) / xx * 100} %>% round(1)
+missPs <- as.matrix(gen$dart_pelo) %>%
+            {xx <<- length(.); . } %>%
+            is.na() %>% {sum(.) / xx * 100} %>% round(1)
+#Number of loci
+# H. molleri
+#  microtellites
+locHm <- adegenet::nLoc(gen$usat_hyla)
+#  SNPs
+locHs <- adegenet::nLoc(gen$dart_hyla)
+# P. cultripes
+#  microtellites
+locPm <- adegenet::nLoc(gen$usat_pelo)
+#  SNPs
+locPs <- adegenet::nLoc(gen$dart_pelo)
+
+#raw raw_genotypes
+genraw <- readRDS("data/intermediate/raw_genotypes.rds")
+rawHm <- adegenet::nInd(genraw$usat_hyla)
+rawHs <- adegenet::nInd(genraw$dart_hyla)
+rawPm <- adegenet::nInd(genraw$usat_pelo)
+rawPs <- adegenet::nInd(genraw$dart_pelo)
+
+#release memory
+rm(genraw, gen)
 
@@
 The above-mentioned sharp clustering with large-scale SNP data compared to microtellites using non-parametric methods does not reflect when using model-based clustering in the popular software STRUCTURE.
-This software provides a powerful tool to estimate individual ancestries from genotypes data (Pritchard et al.
+This software provides a powerful tool to estimate individual ancestries from genotype data (Pritchard et al.
 2000).
+The modeling of allele frequencies can be more powerful (Nassir et al.
+2008), but comes at a high computer cost compared to non-parametric clustering (Porras-Hurtado et al.
+2013), and in many cases, methods such as discriminant analysis of principal components claim to perform better (Jombart et al.
+2008).
 Simulations show equivalent power of tens of microtellites and thousands of SNPs to retrieved with high confidence the correct number of cluster (Haasl and Payseur 2011).
@@
 ## 2.1 Data collection
-We used microsatellite and SNP datasets from two Iberian amphibians, *Hyla molleri* and *Pelobates cultripes*, covering most of their native range (Figure 1).
+We used microsatellite and SNP datasets from two Iberian amphibians, *Hyla molleri* and *Pelobates cultripes*, covering most of their native range (Figure 1; Table S1).
 The microsatellite genotypes were obtained from already-published studies.
-For *Hyla molleri*, we gathered the microsatellite genotypes from *Appendix 1* in Sánchez-Montes et al.
+For *Hyla molleri*, we gathered them from *Appendix 1* in Sánchez-Montes et al.
 (2019).
-They consisted of `r adegenet::nInd(gen$usat_hyla)` samples with `r adegenet::nLoc(gen$usat_hyla)` genotyped microsatellites.
+They consisted of `r rawHm` samples with `r locHm` loci.
 For *Pelobates cultripes*, we requested the genotypes to the authors in Gutiérrez-Rodríguez et al.
 (2017).
-This dataset consisted of `r adegenet::nInd(gen$usat_pelo)` samples with `r adegenet::nLoc(gen$usat_pelo)` genotyped microsatellites.
-The SNP data was generated using DArTseq technology (Diversity Arrays Technology, Canberra, Australia).
-We included samples spanning all over the Iberian Peninsula (Figure 1; Table 1).
-DNA was extracted ExtractMe Genomic DNA 96-Well (DNA GDAŃSK), concentrated with QIAamp DNA Micro (QIAGEN GmbtH) kits, if necessary, and sent for sequencing.
-##
+This dataset consisted of `r rawPm` samples with `r locPm` loci.
+SNP data was generated for `r rawHs` samples of *H.
+molleri* and `r rawPs` samples of *P.
+cultripes*.
+*???for Guillermo: I pasted the following from Gutiérrez-Rogríguez et al.
+(2017).
+Please, check: Most samples were obtained from tail tips of larvae and toes of adults, pre- served in absolute ethanol, and subsequently kept in a freezer at 20 °C???*.
+For *H.
+molleri*, we completed the sampling with tissue from several museum specimens (Table S1).
+Genomic DNA was extracted ExtractMe Genomic DNA 96-Well (DNA GDAŃSK), concentrated with QIAamp DNA Micro (QIAGEN GmbtH) kits, if necessary, and sequenced with DArTseq technology (Diversity Arrays Technology, Canberra, Australia).
 
-## Study area and species
+## 2.2 Data generation
+### 2.2.1 SNPs sequencing
 
-## Data
+Amphibians have large genomes with repetitive regions which hampers genomic sequencing approaches (Funk et al.
+2018; Liedtke et al.
+2018).
+We chose DArTseq, a proprietary protocol to sequence reduced representation of the genome from double-digested fragments.
+DArTseq has been claimed to work well with amphibians because it enriches for high-complexity regions, disregarding repetitive portions of the genome (Lambert et al.
+2016).
+DNA extracts from *H.
+molleri* and *P.
+cultripes* with total DNA ranging from to xxx ug were sent to Diversity Arrays Technology (University of Canberra, Australia).
+The fragments were generated using the restriction enzymes xxx and xxx, and sequenced in an Illumina xxxx as sigle-end reads of 77 nucleotides (nt).
+The sequencing depth for *H.
+molleri* and *P.
+cultripes* were xxx and xxx M reads / sample, respectively.
+Diversity Arrays Technology genotypes the samples using the proprietary DArTSoft14 pipeline and returns a csv file with the genotypes and several parameters on each SNP.
+A xx % of the samples in the run are included as internal replicas to provide confidence levels on the genotype calls.
 
-# 3 | RESULTS
+### 2.2.2 Data filtering
+We additionally applied several filtering steps to the genotype matrices in R 3.5.1 (R Core Team 2018), using functions from the *dartR* 1.1.11 package (Gruber and Georges 2019) and custom functions (https://github.com/csmiguel/usat_snp; Figure S1).
+The genotypes were imported into R and filters were applied following order.
+A per-sample call rate threshold (proportion loci with calls) of 0.35, RepAvg (a repeatability measure of genotypes calls between replicas) threshold of 0.95.
+We used thresholds for allele balance of 0.15 and 0.85, and removed loci for which their coverage was 3.5 times higher than the median coverage (O'Leary et al.
+2018), after which, we filtered loci with a call rate (proportion of samples with a call) below 0.8, retained only one SNP per contig, a set minor allele frequency threshold of 0.02 and removed loci whose alleles deviated from Hardy-Weinberg proportions(O'Leary et al.
+2018).
+To make the datasets more comparable (Figure 1), we filtered out samples from spatially distant localities between marker datasets.
 
-* result 1: sequencing
-We sequenced a total of ´r adegenet::nInd(gen)´ samples which included 1 replica.
-DArTseq genotyped ´r adegenet::nLoc(gen)´SNPs in ´r nlevels(as.factor(gen@other$loc.metrics$CloneID))´ loci.
-The overall missing data was ´r ´ ()
+## 2.3 Population structure
+We conducted model-based genetic structure analyses in STRUCTURE v2.3.4 (Pritchard et al.
+2000).
+For each dataset, we performed 10 replicate runs from K = 1 to K = 8, since the best number of clusters fall within this range for both species (Gutiérrez-Rodríguez et al.
+2017; Sánchez-Montes et al.
+2019).
+We used an admixture model with correlated allele frequencies (Falush et al.
+2003) and no prior information of sample origin.
+For the microtellite data, we used the same run lengths as in their original publications: 5*10^5 burnin followed by 1*10^6 iterations.
+For the SNP datasets, the run lengths were shorter, a burnin of 3*10^4 steps followed by 1 * 10^4 iterations, as for SNP data chain often converge fast (Pritchard, google groups...?).
+For the SNP data, we estimated lambda for K = 1, by averaging lambda estimates across 3 replicate runs.
+The values of lambda (0.60 for *H.
+molleri* and 0.69 for *P.
+cultripes*) were then used across all runs for the SNP data, whereas for microsatellites lambda was fixed to 1.
+These lower values of lambda can improve the modeling of correlated allele frequencies when using SNPs, where often the data is skewed toward rare alleles (Falush et al.
+2003).
+STRUCTURE was run in parallel in 8 cores using Structure_threader (Pina-Martins et al.
+2017), recording steps to log files every 50 and 5000 iterations for the SNP and microsatellite data, respectively.
+We run STRUCTURE analysis with random subsets of loci (n = 200, 500, 1000, 3000, 5000, 10000 and 20000) from the original matrix of SNPs for *P.
+cultripes* and the same STRUCTURE parameters.
+The convergence between the 10 replicate runs for each K was evaluated in R using Gelman and Rubin's convergence diagnostic (GR diagnostic; Gelman and Rubin 1992), with function *gelman.diag* in package *coda* (Plummer et al.
+2006).
 
-* result 2: structure vs dapc
+# 3. RESULTS
+
+After filtering (Figure S1), the dataset kept for analysis included a total `r totalind` individuals from `r nrow(samples)` localities: `r indH` individuals of *Hyla molleri* (microtellites: n = `r indHm`, `r locHm` loci, `r missHm` % missing data; SNPs: n = `r indHs`, `r locHs` loci, `r missHs` % missing data), from `r popH` localities, and `r indP` individuals of *Pelobates cultripes* (microtellites: n = `r indPm`, `r locPm` loci, `r missPm` % missing data; SNPs: n = `r indPs`, `r locPs` loci, `r missPs` % missing data), from `r popP` localities (Figure 1; Table S1).
+
+## 3.1 Model-based clustering
+The convergence of the chains as measured by GR diagnostic varied between speciesa and markers.
+Often, values below 1.05 indicate good convergence (Vats and Knudson 2018 Jeff should check this ref).
+For SNPs, the chains from each K in STRUCTURE converged well up to K = 2 in *H.
+molleri* and up to K = 6 in the case of *P.
+cultripes*.
+Reducing the number of loci did not affect the pattern of convergence across K in a consistent manner (Figures S2.1, S2.2, S2.5 - S2.11; Figure S3.1).
+For microsatellite data, the chains converged consistently up to K = 6 in both species, although the GR diagnostic slightly peaked for K = 5 in *P. cultripes* (Figures S2.3 - S2.4; Figure S3.1).
 

### 2019-07-01 code/01_prepare_data/03_Metadata_consolidation.R

@@ -5,8 +5,7 @@
 # April 2019
 ###.............................................................................
 #GOAL: consolidation of metadata
-#the most.
-#DESCRIPTION:
+#DESCRIPTION: make metadata consitent with indNames
 #PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
 ###.............................................................................
 #  REQUIRED FILES:
@@ -50,7 +49,7 @@ rm(h, i, dart_meta)
 gen <- c(dart_gen, usat_gen)
 rm(dart_gen, usat_gen)
 
-#CONSOLIDATE metadata: removes samples in metadata non present in
+#CONSOLIDATE metadata: removes samples in metadata not present in
 #indNames and reorders metadata$sample_id according to order in indNames
 for (i in seq_along(gen)){
   gen[[i]] <- consolidate_metadata(gen[[i]])

### 2019-07-01 code/03_structure/07_Read_structure_all.R

@@ -15,6 +15,7 @@
 #    Outpath:
 #  DEPENDENCIES:
 ###.............................................................................
+library(dplyr)
 source("code/functions/read_lk_str.r")
 source("code/parameters/structure.r")
 

### 2019-07-01 code/999_plot_convergence_snp_subset.R

@@ -1,33 +1,20 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# April 2019
+###.............................................................................
+#GOAL: Plot convergence only for SNP subsets.
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+
 rhat_results <- readRDS("data/intermediate/rhat.rds")
-conv <- readRDS("data/intermediate/prob_convergence.rds")
 source("code/functions/read_lk_str.r")
 source("code/parameters/structure.r")
 
 
 library(ggplot2)
 library(dplyr)
-
-#Plot convergence
-plot_convergence(method = "with_burnin", p = conv$usat, burnin = burnin_usat)
-plot_convergence(method = "with_burnin", p = conv$dart, burnin = burnin_snp)
-
-#Plot rhat
-p1 <-
-  ggplot(data = rhat_results[[1]], aes(x = k, y = rhat, color = id)) +
-  geom_line() +
-  geom_hline(yintercept = 1.05, col = "red", lty = 2) +
-  coord_cartesian(ylim = c(0.9, 1.2)) +
-  ggtitle(names(rhat_results)[1])
-p2 <-
-  ggplot(data = rhat_results[[2]], aes(x = k, y = rhat, color = id)) +
-  geom_line() +
-  geom_hline(yintercept = 1.05, col = "red", lty = 2) +
-  coord_cartesian(ylim = c(0.9, 1.2)) +
-  ggtitle(names(rhat_results)[2])
-
-pdf(paste0("data/final/rhat.pdf"), height = 6, width = 8)
-  gridExtra::grid.arrange(p1, p2)
-dev.off()
 
 #Plot convergence for marker subsets of Pelobates
 pelo_nloc <-

### 2019-07-01 code/functions/coverage_filt.r

@@ -2,7 +2,7 @@
 # between alleles, units being number of times the most covered allele is
 #different from the other. ie: 2.5 means that the allele with most converage
 #has in average 2.5 times more coverage than the allele with less converage.
-#This step was applied in 10.1111/1755-0998.12997
+#This step was applied in 10.1111/1755-0998.12997 (Tasmanian devil)
 #Thank Elspeth in acknowledgements for sharing part of the code.
 diff_coverage <- function(genlight_object){
   snp <- genlight_object$other$loc.metrics
@@ -32,6 +32,7 @@ include_plot = NULL, plot_name){
   #this function takes genlight formatted with darR::gl.read.dart function
   #and removes unbalanced alleles below and above the tresholds, respectively.
   #it can return plot in working dir. Based on publication 10.1111/mec.14792
+  # O'Leary et al. 2018
   #usage:
   #all_balance_filter(gen, include_plot = T, plot_name = "Allele_balance_filt")
   assertthat::assert_that(class(genlight) == "genlight")

### 2019-07-01 code/functions/read_lk_str.r

@@ -9,6 +9,8 @@ r <- system("find data/final/run_* -name *stlog",
 r
 }
 
+#read log files into a dataframe with:#rowname, path of file#dataset,
+#dataset#k, k# columns with names being the interation and value being the lnlk.
 read_lk <- function(runs = runs){
   p <- runs %>%
     sapply(function(x){
@@ -74,6 +76,7 @@ p = NULL, burnin = NULL){
 }
 
 
+#calculate Gelman and Rubin's convergence diagnostic
 rhat <- function(p = NULL, burnin = NULL){
   it <- as.numeric(names(p)[3:length(names(p))])
   thinning_number <- it[2] - it[1] #frequency of saving steps

### 2019-07-01 code/plot_convergence_allK.R

@@ -1,0 +1,21 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# April 2019
+###.............................................................................
+#GOAL: plot black lines convergence of replicate runs across K for each dataset
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+
+conv <- readRDS("data/intermediate/prob_convergence.rds")
+source("code/functions/read_lk_str.r")
+source("code/parameters/structure.r")
+
+
+library(ggplot2)
+library(dplyr)
+
+#Plot convergence
+plot_convergence(method = "with_burnin", p = conv$usat, burnin = burnin_usat)
+plot_convergence(method = "with_burnin", p = conv$dart, burnin = burnin_snp)

### 2019-07-01 code/plot_cowplots_convergence_tables.R

@@ -1,0 +1,104 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# April 2019
+###.............................................................................
+#GOAL: cowplots convergence Gelman and Rubin's convergence diagnostic
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+
+library(magrittr)
+library(flextable)
+library(grid)
+library(dplyr)
+library(ggplot2)
+
+
+rhat_results <- readRDS("data/intermediate/rhat.rds")
+
+#remove upper_ci column and round numbers
+rhat_results$dart %<>% dplyr::select(-upper_ci) %>%
+  mutate(rhat = round(rhat, digits = 2)) %>%
+  rename(dataset = id)
+rhat_results$usat %<>% dplyr::select(-upper_ci) %>%
+  mutate(rhat = round(rhat, digits = 2)) %>%
+  rename(dataset = id)
+
+
+#replacements in dataset names
+levels(rhat_results$dart$dataset) <-
+  plyr::mapvalues(
+  x = levels(rhat_results$dart$dataset),
+  from = c("1000", "10000", "200", "20000", "3000", "500", "5000",
+  "dart_hyla", "dart_pelo"),
+  to = c("subset 1000", "subset 10000", "subset 200", "subset 20000",
+  "subset 3000", "subset 500", "subset 5000", "H. molleri", "P. cultripes")
+  )
+
+levels(rhat_results$usat$dataset) <-
+  plyr::mapvalues(
+  x = levels(rhat_results$usat$dataset),
+  from = c("usat_hyla", "usat_pelo"),
+  to = c("H. molleri", "P. cultripes")
+  )
+
+#reorder levels
+rhat_results$dart$dataset <-
+factor(rhat_results$dart$dataset,
+   levels = c("H. molleri", "P. cultripes", "subset 200", "subset 500",
+   "subset 1000", "subset 3000", "subset 5000", "subset 10000", "subset 20000"))
+
+#reshape tables
+rusat <-
+  rhat_results$usat %>%
[…]

### 2019-07-02 code/rmarkdown/article.Rmd
kfinder plots

@@
 cultripes* and the same STRUCTURE parameters.
+This approximation allowed to assess the effect of the number of loci in the patterns of genetic structure.
 The convergence between the 10 replicate runs for each K was evaluated in R using Gelman and Rubin's convergence diagnostic (GR diagnostic; Gelman and Rubin 1992), with function *gelman.diag* in package *coda* (Plummer et al.
 2006).
+A central question in population genetics is determining the optimal number of populations subjacent to the genetic structure (Allendorf and Luikart 2007).
+However, hierarchical population structure (Evanno et al.
+2005), unbalanced sampling and number markers (Wang 2017) can lead to the incorrect inference of number of populations.
+We used KFinder (Wang 2019) to compare the best number of populations for each dataset through 3 approaches: (1) Pr[X|K], the probability of data X given K populations (Pritchard et al.
+2000) and, (2) ΔK, the most popular method and which considers the rate of change in the log probability of data between successive K (Evanno et al.
+2005) and 3) PI, parsimony index, a newly proposed metric which favors K's which yield most consistent and with minimal average individual admixture, and which claims to be more consistent across wider demographic scenarios (Wang 2019).
 
@@
 ## 3.1 Model-based clustering
-The convergence of the chains as measured by GR diagnostic varied between speciesa and markers.
+Convergence of the chains measured by the GR diagnostic varied between species and markers.
 Often, values below 1.05 indicate good convergence (Vats and Knudson 2018 Jeff should check this ref).

### 2019-07-02 code/03_structure/09_readKfinder.R

@@ -4,7 +4,7 @@
 # https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
 # April 2019
 ###.............................................................................
-#GOAL: calculate best K
+#GOAL: read kfinder results r
 #PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
 ###.............................................................................
 #  REQUIRED FILES:
@@ -17,7 +17,6 @@
 ###.............................................................................
 library(dplyr)
 library(magrittr)
-library(dplyr)
 #1. create parameters file
 # vector with folder names for each run
 kfinder <- dir("data/final", pattern = "str.K$", full.names = T, recursive = T)
@@ -35,44 +34,6 @@ names(kfinder_ls) <- kfinder %>%
   stringr::str_extract("run.*/") %>%
   gsub(pattern = "/", replacement = "")
 
-#plot parsimony
-plot_k <- function(klist = NULL, index = NULL, runs_pattern = NULL, mode = 1){
-  #runs_pattern, is the pattern to grep from names
-  #index is Pritchard, Evanno or Parsimony
-  #list with data frames of k statistics
-  to_plot <- klist[grep(runs_pattern, names(klist))]
-  if (mode == "subsampling"){
-    to_plot <-
-      to_plot[names(to_plot) %>%
-                stringr::str_extract("[0-9]+") %>% as.numeric() %>% order]
-  }
-  y_lim <- to_plot %>% sapply(function(x) x[[index]]) %>%
-    unlist %>% as.numeric() %>% range(na.rm = T)
-  plot(1, type = "n",
-       xlim = range(to_plot[[1]]$K),
-       ylim = y_lim,
-       ylab = index,
-       xlab = "K")
-  cols <- colorRampPalette(c("blue", "red"))
-  for (i in seq_along(to_plot)){
-    lines(to_plot[[i]]$K,
-          as.numeric(as.character(to_plot[[i]][[index]])),
-          lwd = 2,
-          type = "o",
-          col = cols(length(to_plot))[i]
-          )
-  }
-  legend("bottomright", legend = names(to_plot),
-         col = cols(length(to_plot)), pch = 1, lwd = 2,
-         bg = adjustcolor("white", alpha.f = 0.5))
-}
-plot_k(klist = kfinder_ls, index = "Parsimony", runs_pattern = "pelo")
[…]

### 2019-07-02 code/functions/plot_kfinder.r

@@ -1,0 +1,41 @@
+library(dplyr)
+
+#function to plot result from Kfinder saved in kfinder_ls.rds
+
+plot_k <- function(klist = NULL, index = NULL, runs_pattern = NULL, mode = 1){
+  #runs_pattern, is the pattern to grep from names
+  #index is Pritchard, Evanno or Parsimony
+  #list with data frames of k statistics
+  to_plot <- klist[grep(runs_pattern, names(klist))]
+  if (mode == "subsampling"){
+    to_plot <-
+      to_plot[names(to_plot) %>%
+                stringr::str_extract("[0-9]+") %>% as.numeric() %>% order]
+  }
+  y_lim <- to_plot %>% sapply(function(x) x[[index]]) %>%
+    unlist %>% as.numeric() %>% range(na.rm = T)
+
+  ylabs <- plyr::mapvalues(
+    x = index,
+    from = names(klist[[1]])[-1],
+    to = c("Pr[X|K]", "Evanno ΔK", "Parsimony index")
+  )
+
+  plot(1, type = "n",
+       xlim = range(to_plot[[1]]$K),
+       ylim = y_lim,
+       ylab = ylabs,
+       xlab = "K")
+  cols <- colorRampPalette(c("#d73027", "#1a9850"))
+  for (i in seq_along(to_plot)){
+    lines(to_plot[[i]]$K,
+          as.numeric(as.character(to_plot[[i]][[index]])),
+          lwd = 2,
+          type = "o",
+          col = cols(length(to_plot))[i]
+          )
+  }
+  legend("bottomright", legend = names(to_plot),
+         col = cols(length(to_plot)), pch = 1, lwd = 2,
+         bg = adjustcolor("white", alpha.f = 0.5))
+}

### 2019-07-02 code/plot_Kfinder.R

@@ -1,0 +1,38 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# April 2019
+###.............................................................................
+#GOAL: plot
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+library(dplyr)
+
+#load kfinder results
+kfinder_ls <- readRDS("data/intermediate/kfinder_ls.rds")
+source("code/functions/plot_kfinder.r")
+
+#replace names
+names(kfinder_ls) <-
+  plyr::mapvalues(
+    x = names(kfinder_ls),
+    from = c("run_1000", "run_10000", "run_200", "run_20000", "run_3000",
+      "run_500", "run_5000", "run_dart_hyla", "run_dart_pelo",
+      "run_usat_hyla", "run_usat_pelo"),
+    to = c("subset 1000", "subset 10000", "subset 200", "subset 20000",
+           "subset 3000", "subset 500", "subset 5000", "SNPs H. molleri",
+           "SNPs P. cultripes", "microsatellites H. molleri",
+           "microsatellites P. cultripes"))
+
+#plot parsimony
+
+pdf("data/final/kfinder_subsampling.pdf", height = 9, width = 13)
+  par(mfrow = c(3, 3))
+  for (i in names(kfinder_ls[[1]])[-1]){
+    plot_k(klist = kfinder_ls, index = i,
+      runs_pattern = "[0-9]", mode = "subsampling")
+    plot_k(klist = kfinder_ls, index = i, runs_pattern = "cult")
+    plot_k(klist = kfinder_ls, index = i, runs_pattern = "molle")
+  }
+dev.off()

### 2019-07-02 code/plot_structure.R

@@ -1,0 +1,37 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# April 2019
+###.............................................................................
+#GOAL: plot
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+library(ggplot2)
+
+#read structure output
+h <- pophelper::readQStructure(best_files[1], indlabfromfile = T)
+
+#sort indlabels by populations
+# read genotypes
+gen <- readRDS("data/intermediate/gen_consolidated_filtered.rds")
+pelo <- gen$dart_pelo$other$metadata %>%
+          dplyr::arrange(locality)
+
+#sort ancestry table according to order of populations
+h$str_K6_rep3_f %<>%
+  .[match(rownames(.), pelo$sample_id), ]
+#plot structure output
+p <-
+  pophelper::plotQ(h,
+                   exportplot = F,returnplot = T,
+                   showindlab = T, indlabsize = 8,
+                   height = 3, barsize = 1, useindlab = T,
+                   showtitle = T, barbordercol = "white",
+                   showyaxis = T, sortind = "label",
+                   ordergrp=T, grpmean = F, grplabsize=2,
+                   grplabangle = 60,linepos = 1,
+                   grplabpos=0.9,splabsize = 8,
+                   grplabheight = 3)
+
+plot(p$plot[[1]])

### 2019-07-09 code/rmarkdown/article.Rmd
clumpak

@@
 2016; Wang 2016).
-Neighbor joining clustering is a complementary way to overcome some limitations of model-based clustering and to recover a reliable evolutionary history of the populations under a wide variaty of scenarios (Kalinowski 2009, 2011).
-The reconstructions of phylogenetic trees using neighbor joining with microtellite data are often coherent with the phylogeographic structure recovered with SNPs.
+Neighbor joining clustering is a complementary way to overcome some limitations of model-based clustering and to recover a reliable evolutionary history of the populations under a wide variety of scenarios (Kalinowski 2009, 2011).
+The reconstruction of phylogenetic trees using neighbor joining with microtellite data are often coherent with the phylogeographic structure recovered with SNPs.
 However, the large homoplasy undermines their power to retrieve phylogenetic relationships between distant populations (Estoup et al.
@@
 2012; Bradbury et al.
-2018), although when several hundreds of loci are used the correct topology is recovered correctly (Takezaki and Nei 2006).
+2018), although when several hundreds of loci are used the correct topology is recovered correctly (Takezaki and Nei 2008).
 The large amount of SNPs and their genome-wide distribution secure a range mutation rates which can provide sufficient information at different evolutionary scales, from recent within species processes to interspecies phylogenies (eg.
@@
 
+## 2.4 Phylogenetic structure
+
+We evaluated the phylogenetic relationships between the populations using the neighbor-joining (NJ) algorithm (Saitou and Nei 1987).
+We discarded using model-based phylogenetic inference for the implicit difficulties of fitting evolutionary and demographic models for intra-species trees (), and for the possibilities that offer comparing the 2 marker types (SNPs and microtellites) under a similar tree-reconstruction method.
+The application of NJ relies on a distance matrix computed from the genotypes.
+For the microtellite datasets we computed these matrices using Da distance (Nei et al.
+1983) with *genet.dist* function from *hierfstat* 0.04-22 package (Goudet and Jombart 2015) in R.
+Da distances have been shown to yield more reliable topologies to other distances (Takezaki and Nei 1996, 2008).
+For the SNP genotypes, we computed euclidean distances with *stats::dist* function in R.
+Then NJ trees were reconstructed with *njs* function from *ape* 5.0 (Paradis and Schliep 2018) for each dataset.
+We computed the bootstrap support with a modified approach from *ape::boot.phylo* to accommodate non-standard input data.
+First, we resampled with replacement loci in the genotype matrices, from which we recomputed NJ trees.
+These trees were rooted using midpoint rooting with function *midpoint.root* from *phytools* 0.6-60 (Revell 2012) in R.
+And finally, bootstrap values were computed by counting the partitions of the original trees present in 1000 bootstrapped trees.
+*Marker type and bootstrap support*.
+For each tree, branch lengths were normalized by dividing edge lengths by the length between the most distant tips in the tree.
+Then, we used several functions in *treeman* 1.1.3 package (Bennett 2018) in R, to extract 2 metrics from each node in the trees: 1) the distance from each node to the root (Dnr), and 2) the distance from each node to its parent node (Dnp).
+??For the nodes in each tree we assessed the relation of Dnr and Dnp against their corresponding bootstrap support values fitting generalized linear models (GMLs) using a logit link function, with *stats::glm* function in R.
+*Actually the fit is really bad.
+See script 99_04_model_fitting.R and ask Jeff*??
+
+*Number of markers and bootstrap support*.
+
+*MAF and bootstrap support*.
+
 # 3. RESULTS
@@
 ## 3.1 Model-based clustering
+*Convergence*.
 Convergence of the chains measured by the GR diagnostic varied between species and markers.
@@
 For microsatellite data, the chains converged consistently up to K = 6 in both species, although the GR diagnostic slightly peaked for K = 5 in *P. cultripes* (Figures S2.3 - S2.4; Figure S3.1).
+*Optimal K*.
+The most optimal number of populations was evaluated with 3 metrics.
+Pr[X|K] increased steadily from K = 2 to K = 8 in all datasets (Figure 2 A-C).
+Evanno's ΔK peaked at K = 2 in all SNP datasets, followed by K = 3 (Figure 2 D-F).
+However, in the microsatellite dataset from *H.
+molleri*, ΔK peaked at K = 4 and K = 6 (Figure 2D).
+The predominant K = 2 in the SNP datasets were not in agreement with results from the Parsimony index (Figure 2 G-I), except for the microtellite data in *H.
+molleri*, in which PI and ΔK peaked at K = 4 (Figure 2G).
+For SNPs in *H.
+molleri*, the highest values of the PI were at K = 4 and K = 5, contrary to the K = 2 from ΔK (Figure 2G).
+In *P.
+cultripes*, the PI peaked at K = 4 for both markers, and then sunk in the SNP data, whereas for  microtellites the PI reached its maxima at K = 6 - K = 8 (Figure 2H).
+In the subsets of SNPs from *P.
+cultripes* the PI increased gradually up to K = 4.
+After that, the distinct marker sets showed inconsistent trends (Figure 2I).
 
-* result 3: genetic diversity
+## 3.2 Phylogenetic structure
+
+*Coherence between model-based and NJ clustering*.
+
+*Marker type and bootstrap support*.
+
+*Number of markers and bootstrap support*.
+
+*MAF and bootstrap support*.
 
@@
 * Key result 2
+The little bootstrap support we describe in phylogenetic trees from microtellite data is not surprising, given that often several tens of microsatellites or more are needed to recover the correct topology of trees (Nei et al.
+1983; Takezaki and Nei 1996)
 

### 2019-07-09 code/02_phylogenies/02_BS_analysis.R

@@ -28,18 +28,15 @@ bt <- readRDS(file = input_path2)
 source("code/functions/bs_analysis.r")
 source("code/parameters/boot.r")
 
+#get list of dataframes with columns being:
+# dnr, distance from node to root
+# dnp, distance from node to parent node
+# bs_tm, bootstrap support
+#and rows being nodes.
+
 bs_treemetrics <-
   seq_along(tr) %>%
     lapply(function(x) tree_metrics(tr[[x]], bt[[x]]))
 names(bs_treemetrics) <- names(tr)
 
 saveRDS(bs_treemetrics, "data/intermediate/bs_treemetrics.rds")
-  #statistics
-  dnnmin.glm <- summary(glm(bs~dnnmin, family = binomial()))
-    dnnmin.glm.e <- dnnmin.glm$coefficients[2, 1] %>% round(3)
-    dnnmin.glm.s <- dnnmin.glm$coefficients[2, 4] %>%
-      formatC(format = "e", digits = 2)
-  dno.glm <- summary(glm(bs~dno, family = binomial()))
-    dno.glm.e <- dno.glm$coefficients[2, 1] %>% round(3)
-    dno.glm.s <- dno.glm$coefficients[2, 4] %>%
-      formatC(format = "e", digits = 2)

### 2019-07-09 code/02_phylogenies/99_04_model_fitting.R

@@ -1,0 +1,33 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# April 2019
+###.............................................................................
+#GOAL: fit models to bootstrap data
+#DESCRIPTION:
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+library(dplyr)
+
+bs_treemetrics <- readRDS <- ("data/intermediate/bs_treemetrics.rds")
+#i tried to fit glm with binomial distirbution and others but they do not fit well the data.
+m1 <- lm(bs_tm~log(dnp), data = bs_treemetrics$dart_pelo)
+m2 <- glm(bs_tm~dnp*dnr, data = bs_treemetrics$dart_pelo, family = "binomial")
+
+summary(m1)
+plot(bs_tm~dnp, data = bs_treemetrics$dart_pelo)
+
+#also gam do not fit well:
+
+#http://environmentalcomputing.net/intro-to-gams/
+library(mgcv)
+gam_y <- mgcv::gam(bs_tm ~ s(dnp), data = h, method = "REML")
+y_pred <- predict(gam_y, data.frame(dnp = seq(0, max(h$dnr), length.out = 1000)))
+
+ggplot(data.frame(h), aes(dnp, bs_tm)) +
+  geom_point() +
+  geom_smooth(method = gam, formula = y ~ s(x))
+
+gam.check(gam_y)
+summary(gam_y)

### 2019-07-09 code/03_structure/10_create_clumpak_input.sh

@@ -1,0 +1,42 @@
+#!/bin/bash
+#put all your results in one folder ‘results_folder’
+#Go to the command line terminal and type:
+#1. prepare input file. It needs the below structure:
+# Structure results zipped into a single file
+# stresults.zip
+# 	k1.zip
+#		k1/
+#			project_data_k1_run10_f
+#			project_data_k1_run1_f
+#			...
+#	k2.zip
+#		k2/
+#			...
+#1. 1. create folder STRUCTURE
+rm -rf mkdir data/intermediate/clumpak
+mkdir data/intermediate/clumpak
+
+find data/final -name run_* | while read dataset
+#loop 1: iteration across datasets
+do
+d=$(echo $dataset | sed 's/^.*\///g')
+mkdir data/intermediate/clumpak/$d
+  for k in {1..8}
+  #loop 2: iteration across K
+  do
+  outdir=$(echo data/intermediate/clumpak/$d/K$k)
+  mkdir $outdir
+  #list with runs within for Kx within dataset
+    find $dataset -name *K"$k"_*_f | while read str_run
+    #loop 3: iteration across runs within K
+    do
+      #store repetition id in variable
+    rep=$(echo $str_run | sed 's/.*\(rep\)/\1/')
+      #create output name for str_run
+    new=$(echo $outdir/K"$k"_"$rep")
+    cp $str_run $new
+    done
+  done
+  cd data/intermediate/clumpak/ && zip -r "$d".zip "$d" -x "*.DS_Store" && cd -
+  rm -r data/intermediate/clumpak/$d
+done

### 2019-07-09 code/11__clumpak_main_pipeline.sh

@@ -1,0 +1,9 @@
+#main pipeline
+#/Applications/CLUMPAK/ should contain downloaded folders from http://clumpak.tau.ac.il/
+find data/intermediate/clumpak -name *.zip | while read str_zipped
+do
+outdir=(echo $str_zipped | sed 's/^.*\///g')
+perl /Applications/CLUMPAK/CLUMPAK.pl --id 1 \
+--dir data/intermediate/clumpak/"$outdir" \
+--file str_zipped
+done

### 2019-07-09 code/99-plot_BS_analysis.R

@@ -1,26 +1,0 @@
-#plot results from bootstrap analysis
-bstree <- readRDS("data/intermediate/bs_treemetrics.rds")
-# plot parameters
-x_lab <- c("Distance to parent node", "Distance from root")
-max_dnr <- sapply(bstree, function(x) x$dnr) %>% unlist %>% max
-max_dnp <- sapply(bstree, function(x) x$dnp) %>% unlist %>% max
-x_lim <- matrix(c(0, max_dnp, 0, max_dnr), ncol = 2)
-
-#produce plots
-for (i in 1:2){
-  pdf(file = paste0("data/final/",
-  gsub(" ", "_", x_lab[i]), ".pdf"),
-   height = 8, width = 10)
-  par(mfrow = c(2, 2))
-    for (p in 1:length(bstree)){
-      hh <- bstree[[p]]
-      plot(as.numeric(as.matrix(hh[, i])),
-          hh$bs_tm, pch = 20,
-          xlab = x_lab[i],
-          ylab = "Bootstrap support",
-          main = names(bstree)[p],
-          xlim = x_lim[, i],
-          ylim = c(0, 1))
-      }
-  dev.off()
-}

### 2019-07-09 code/99-plot_trees.R

@@ -1,20 +1,0 @@
-bt <- readRDS("data/intermediate/boot_support.rds")
-tr <- readRDS("data/intermediate/bt_trees.rds")
-
-#plot trees
-  nodecol <- bt
-for (i in seq_along(names(bt))){
-  nodecol[[i]][nodecol[[i]] < tresh * nboot] <- "red"#color nodes bt BS
-  nodecol[[i]][nodecol[[i]] != "red"] <- "black"
-  pdf(file = paste0("data/final/", names(bt)[[i]], "bs_tree.pdf"))
-  plot(tr[[i]][[1]], "u", #plot fan
-       use.edge.length = T,
-       show.tip.label = T, cex = 0.4,
-       edge.color = "black")
-  nodelabels(pch = 20, col = nodecol[[i]], cex = 0.5) #issue assert that length of color nodes is the same as the number of nodes inf the tree
-  add.scale.bar()
-  title(paste0("NJ tree ", names(bt)[i], ", ", nLoc(gen[[i]]), " loci, ", nboot,
-        " bootstrap")) #title plot
-  dev.off()
-      }
-rm(i, tresh)

### 2019-07-09 code/Figure3_markertype-bs.R

@@ -1,0 +1,43 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# June 2019
+###.............................................................................
+#GOAL: Plot results from bootstrap analysis
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+#plot results from bootstrap analysis
+bstree <- readRDS("data/intermediate/bs_treemetrics.rds")
+# plot parameters
+x_lab <- c("Distance to parent node", "Distance from root")
+max_dnr <- sapply(bstree, function(x) x$dnr) %>% unlist %>% max
+max_dnp <- sapply(bstree, function(x) x$dnp) %>% unlist %>% max
+x_lim <- matrix(c(0, max_dnp, 0, max_dnr), ncol = 2)
+
+#rename datasets
+bstree
+names(bstree) <-
+  plyr::mapvalues(
+    x = names(bstree),
+    from = names(bstree),
+    to = c("SNPs H. molleri", "SNPs P. cultripes",
+    "microtellites H. molleri", "microtellites P. cultripes"))
+#produce plots
+for (i in 1:2){
+  pdf(file = paste0("data/final/",
+  gsub(" ", "_", x_lab[i]), ".pdf"),
+   height = 7, width = 8.8)
+  par(mfrow = c(2, 2), las = 1)
+    for (p in 1:length(bstree)){
+      hh <- bstree[[p]]
+      plot(as.numeric(as.matrix(hh[, i])),
+          hh$bs_tm, pch = 20,
+          xlab = x_lab[i],
+          ylab = "Bootstrap support",
+          main = names(bstree)[p],
+          xlim = x_lim[, i],
+          ylim = c(0, 1))
+      }
+  dev.off()
+}

### 2019-07-09 code/Figurex_plot_trees.R

@@ -1,0 +1,63 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# June 2019
+###.............................................................................
+#GOAL: Plot results from bootstrap analysis
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+library(ape)
+library(dplyr)
+
+#bootstrap support
+bt <- readRDS("data/intermediate/boot_support.rds")
+#trees
+tr <- readRDS("data/intermediate/bt_trees.rds")
+#genotypes
+gen <- readRDS("data/intermediate/gen_consolidated_filtered.rds")
+#best files as determined by KFinder
+best_files <- readRDS("data/intermediate/best_files_kfinder.rds")
+
+source("code/parameters/boot.r")
+
+#rename datasets
+names(tr) <-
+  plyr::mapvalues(
+    x = names(tr),
+    from = names(tr),
+    to = c("SNPs H. molleri", "SNPs P. cultripes",
+    "microtellites H. molleri", "microtellites P. cultripes"))
+
+#plot trees
+plot_nj <-  function(bt = bt, thresh = thresh, coln = c("red", "black"),
+                tips = c("names", "ancestry"), output = "bs_tree.pdf"){
+  #bt is vector with bootstrap supports
+  #thresh is the threshold to color nodes
+  #coln are the colors for below, above threshold
+  #tips, names: show individual labels, ancestry: shows pie with ancestry.
+  nodecol <- bt
+  if(tips == "names") tip <- T else tip <- F
+  for (i in seq_along(bt)){
+    nodecol[[i]][nodecol[[i]] < thresh * nboot] <- coln[1]#color nodes bt BS
+    nodecol[[i]][nodecol[[i]] != coln[1]] <- coln[2]
+    pdf(file = paste0("data/final/", names(bt)[[i]], output))
+    plot(tr[[i]][[1]], "u", #plot fan
+         use.edge.length = T,
+         show.tip.label = tip, cex = 0.4,
+         edge.color = "grey")
+    #issue assert that length of color nodes is the same as the number of nodes inf the tree
[…]

### 2019-07-09 code/parameters/boot.r

@@ -1,5 +1,5 @@
-nboot = 500 #number of bootstrap
-tresh = .7 #treshold for coloring nodes
+nboot = 1000 #number of bootstrap
+thresh = .7 #treshold for coloring nodes
 #subsample sizes for pelobates dartseq data
 s <- c(25000, 20000, 15000, 10000, 5000, 3000, 1000, 500, 200)
 #times the markers are subsampled from the original matrix of genotypes

### 2019-07-09 code/plot_structure.R

@@ -8,7 +8,7 @@
 #PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
 ###.............................................................................
 library(ggplot2)
-
+best_files <- readRDS("data/intermediate/best_files_kfinder.rds")
 #read structure output
 h <- pophelper::readQStructure(best_files[1], indlabfromfile = T)
 

### 2019-07-09 code/02_structure/11_pop_files_clumpak.R
run clumpak

@@ -1,0 +1,41 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# April 2019
+###.............................................................................
+#GOAL: create podtopop populations_file for CLUMPAK
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+library(dplyr)
+library(dartR)
+
+input_path1 <- "data/intermediate/gen_consolidated_filtered.rds"
+input_path2 <- "data/intermediate/table1.rds"
+gen <- readRDS(file = input_path1)
+meta <- readRDS(file = input_path2)
+
+#structure runs datasets
+str_files <- dir(path = "data/final", pattern = "run_")
+
+#names from gen object that correspond to the datset
+gen_names <-
+  str_files %>% replace(grep("[0-9]", str_files), "run_dart_pelo") %>%
+  gsub(pattern = "run_", replacement = "")
+
+index_pop <- cbind(str_files, gen_names)
+
+#for each dataset create a pop file
+seq_along(str_files) %>%
+lapply(function(x){
+  #select element from list that matches names
+  subset_gen <- gen[grep(index_pop[x, 2], names(gen))][[1]]
+  #create table with IDs
+  subset_gen$other$metadata %>%
+    dplyr::left_join(meta, by = c("locality" = "locality")) %>%
+    as.data.frame %>%
+    dplyr::select(ID) %>%
+    write.table(file = file.path("data/intermediate/clumpak",
+                         paste0(index_pop[x, 1], "_populations_file")),
+                col.names = F, quote = F, row.names = F)
+})

### 2019-07-09 code/02_structure/12__clumpak_main_pipeline.sh

@@ -1,0 +1,18 @@
+#!/bin/bash
+#main pipeline in CLUMPAK
+#/Applications/CLUMPAK/ should contain downloaded folders from http://clumpak.tau.ac.il/
+#I have to run CLUMPAK from the app path instead of my project folder because CLUMPAK calls
+#other scripts within the app folder from relative paths
+h=$(pwd)
+cd /Applications/CLUMPAK
+
+find "$h"/data/intermediate/clumpak -name *.zip | while read str_zipped
+do
+dataset=$(echo $str_zipped | sed 's/^.*\///g' | sed 's/.zip//')
+pop_file=$(echo "$dataset"_populations_file)
+perl CLUMPAK.pl --id 1 \
+  --dir "$h"/data/intermediate/clumpak/$dataset \
+  --file $str_zipped \
+  --podtopop "$h"/data/intermediate/clumpak/$pop_file
+done
+cd -

### 2019-07-09 code/11__clumpak_main_pipeline.sh

@@ -1,9 +1,0 @@
-#main pipeline
-#/Applications/CLUMPAK/ should contain downloaded folders from http://clumpak.tau.ac.il/
-find data/intermediate/clumpak -name *.zip | while read str_zipped
-do
-outdir=(echo $str_zipped | sed 's/^.*\///g')
-perl /Applications/CLUMPAK/CLUMPAK.pl --id 1 \
---dir data/intermediate/clumpak/"$outdir" \
---file str_zipped
-done

### 2019-07-09 code/02_structure/13_collect_clumpak_major.R
plot trees

@@ -1,0 +1,37 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# April 2019
+###.............................................................................
+#GOAL: collect clumpak output for major cluster
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+library(dplyr)
+#Collect Clumpak output:
+datasets <-
+  dir("data/intermediate/clumpak", pattern = "run_", full.names = T) %>%
+    {.[-grep(., pattern = ".zip|_file")]}
+
+#create list with Major Cluster ancestries
+qclumpak <-
+  seq_along(datasets) %>%
+  lapply(function(x){
+    #kruns <- dir(path = datasets[x], pattern = "K=[0-9]")
+    k <- list()
+    for (i in 1:8){
+      k[[i]] <-
+        paste0(datasets[x], "/K=", i,
+          "/MajorCluster/CLUMPP.files/ClumppIndFile.output") %>%
+        readLines() %>%
+        gsub(pattern = "^.*: ", replacement = "") %>%
+        textConnection %>%
+        read.table
+      names(k)[i] <- paste0("K", i)
+    }
+    k
+  })
+
+names(qclumpak) <- datasets %>% gsub(pattern = "^.*\\/run_", replacement = "")
+
+saveRDS(qclumpak, "data/intermediate/clumpak_major.rds")

### 2019-07-09 code/Figurex_plot_trees.R

@@ -9,6 +9,7 @@
 ###.............................................................................
 library(ape)
 library(dplyr)
+library(adegenet)
 
 #bootstrap support
 bt <- readRDS("data/intermediate/boot_support.rds")
@@ -18,9 +19,13 @@ tr <- readRDS("data/intermediate/bt_trees.rds")
 gen <- readRDS("data/intermediate/gen_consolidated_filtered.rds")
 #best files as determined by KFinder
 best_files <- readRDS("data/intermediate/best_files_kfinder.rds")
+#clumpak output
+qclumpak <- readRDS("data/intermediate/clumpak_major.rds")
+#pop ids
+meta <- readRDS("data/intermediate/table1.rds")
 
 source("code/parameters/boot.r")
-
+source("code/functions/plot_nj_trees.r")
 #rename datasets
 names(tr) <-
   plyr::mapvalues(
@@ -30,34 +35,11 @@ names(tr) <-
     "microtellites H. molleri", "microtellites P. cultripes"))
 
 #plot trees
-plot_nj <-  function(bt = bt, thresh = thresh, coln = c("red", "black"),
-                tips = c("names", "ancestry"), output = "bs_tree.pdf"){
-  #bt is vector with bootstrap supports
-  #thresh is the threshold to color nodes
-  #coln are the colors for below, above threshold
-  #tips, names: show individual labels, ancestry: shows pie with ancestry.
-  nodecol <- bt
-  if(tips == "names") tip <- T else tip <- F
-  for (i in seq_along(bt)){
-    nodecol[[i]][nodecol[[i]] < thresh * nboot] <- coln[1]#color nodes bt BS
-    nodecol[[i]][nodecol[[i]] != coln[1]] <- coln[2]
-    pdf(file = paste0("data/final/", names(bt)[[i]], output))
-    plot(tr[[i]][[1]], "u", #plot fan
-         use.edge.length = T,
-         show.tip.label = tip, cex = 0.4,
-         edge.color = "grey")
-    #issue assert that length of color nodes is the same as the number of nodes inf the tree
-    nodelabels(pch = 20, col = nodecol[[i]], cex = 0.5)
-    if (tips == "ancestry"){
-      h <- pophelper::readQStructure(best_files[8], indlabfromfile = T)[[1]] %>%
-       as.matrix
[…]

### 2019-07-09 code/functions/plot_nj_trees.r

@@ -1,0 +1,60 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# April 2019
+###.............................................................................
+#GOAL: plot nj trees
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+
+plot_nj <-  function(bt = bt, thresh = thresh, coln = c("red", "black"),
+                tips = c("names", "ancestry"), typeP = c("phylogram", "unrooted"),
+                output = "bs_tree.pdf", pchn = c(20, 18), K = "K2"){
+  #bt is vector with bootstrap supports
+  #thresh is the threshold to color nodes
+  #coln are the colors for below, above threshold
+  #tips, names: show individual labels, ancestry: shows pie with ancestry.
+  nodecol <- bt
+  if(tips == "names") tip <- T else tip <- F
+  for (i in seq_along(bt)){
+    nodecol[[i]][nodecol[[i]] < thresh * nboot] <- coln[1]#color nodes bt BS
+    nodecol[[i]][nodecol[[i]] != coln[1]] <- coln[2]
+    pchnod <- plyr::mapvalues(nodecol[[i]], coln, pchn) %>% as.numeric
+    pdf(file = paste0("data/final/",
+      paste(names(bt)[[i]], K, typeP, output, sep = "_")))
+    plot(tr[[i]][[1]], type = typeP,
+         use.edge.length = T,
+         show.tip.label = tip, cex = 0.4,
+         edge.color = "grey")
+    #issue assert that length of color nodes is the same as the number of nodes inf the tree
+    ape::nodelabels(pch = pchnod, col = nodecol[[i]], cex = 0.8)
+    add.scale.bar()
+    title(paste0("NJ tree ", names(tr)[i], ", ", adegenet::nLoc(gen[[i]]),
+                 " loci, ", nboot, " bootstrap"), cex.main = 0.7) #title plot
+    if (tips == "ancestry"){
+      #get ancestries from clumpak object
+      h <- qclumpak[[names(gen)[i]]][[K]] %>%
+        as.matrix %>%
+        `rownames<-`(adegenet::indNames(gen[[names(gen)[i]]]))
[…]

### 2019-07-10 code/rmarkdown/Figures.rmd
subsample trees

@@
 ```
+##Figure 2:
+
+Assessment of optimal K values using (A-C) Pr[X|K] (Pritchard et al.
+2000), (D-F) ΔK (Evanno et al.
+2005) and (G-I) the parsimony index, PI (Wang 2019), from K = 1 to K = 8, for *H.
+molleri* (A, D, G), *P.
+cultripes* (B, E, H) and their subsampled datasets from n loci (n in 200, 500, 1000, 3000, 5000, 10000, 20000) (C, F, I).
+Panels A, B, D, E, G and H, depict results from SNP (red) and microsatellite (green) data, whereas panels C, F and I, depict results from SNP data alone.??I need to overplot SDs, change the line and pch pattern for usat/snp and color palette for subsets??
+
+##Figure 3
+
+Relation of bootstrap support to Dnr (distance from node to root) and Dnp (distance from node to parent node), for SNP (A, B, E, F) and microtellite (C, D, G, H) data from *H.
+molleri* (A, C, E, G) and *P.
+cultripes* (B, D, F, H).??should I overplot here the results from loci subsampling??
+
+##Figure 4
+
+Rooted neighbor joining trees for SNP (A, B) and microtellite (C-D) datasets from *H.
+molleri* (A, C) and *P.
+cultripes* (B, D).
+Nodes with bootstrap support above 0.70 are represented with black diamonds, or else with red dots.
+In each tip, pie chart depicts the ancestry proportions from CLUMPAK consensus solutions for the major cluster for K = 4 together with the population ID from Table 1.

### 2019-07-10 code/rmarkdown/article.Rmd
subsample trees

@@
 2005) and 3) PI, parsimony index, a newly proposed metric which favors K's which yield most consistent and with minimal average individual admixture, and which claims to be more consistent across wider demographic scenarios (Wang 2019).
+We ran CLUMPAK (Kopelman et al.
+2015) on STRUCTURE outputs.
+CLUMPAK feeds the software CLUMPP with the replicate runs at each value of K to generate consensus solutions for the distinct modes.
+It also computes the similarity between ancestry matrices from each run and it matches clusters across successive values of K.
 
@@
 And finally, bootstrap values were computed by counting the partitions of the original trees present in 1000 bootstrapped trees.
+We inspected the phylogeographic congruence between STRUCTURE and the phylogenetic by plotting together a the consensus solution for the major cluster from the output of CLUMPAK for K = 4, since K = 4, seemed to be an optimal solution for many of the datasets (Figure 2).
 *Marker type and bootstrap support*.
@@
 # 3. RESULTS
-
 After filtering (Figure S1), the dataset kept for analysis included a total `r totalind` individuals from `r nrow(samples)` localities: `r indH` individuals of *Hyla molleri* (microtellites: n = `r indHm`, `r locHm` loci, `r missHm` % missing data; SNPs: n = `r indHs`, `r locHs` loci, `r missHs` % missing data), from `r popH` localities, and `r indP` individuals of *Pelobates cultripes* (microtellites: n = `r indPm`, `r locPm` loci, `r missPm` % missing data; SNPs: n = `r indPs`, `r locPs` loci, `r missPs` % missing data), from `r popP` localities (Figure 1; Table S1).
@@
 After that, the distinct marker sets showed inconsistent trends (Figure 2I).
+*clumpak table*
 
 ## 3.2 Phylogenetic structure
-
 *Coherence between model-based and NJ clustering*.
+The phylogeographic structure of the model-based ancestries for the SNP dataset were coherent with the topology of their phylogenetic trees (Figure 4 A-B).
+That is, the samples from the same populations clustered together on the same or close clades and the ancestries from the samples were coherent with the phylogenetic relationships at different depths in the tree.
+However, the topology of the trees for microtellites were little supported.
+Furthermore, it neither provided a solid grouping of the samples from the same populations and it was not able to explain the relationships at different depths in the trees (Figure 4 C-D).
 
 *Marker type and bootstrap support*.
+The topology of the trees from SNP data were highly supported compared to the trees from microtellites (Figures 3 and 4).
+The overall support was higher for the SNP dataset of *P.
+cultripes* than for *H.
+molleri* (Figure 3 A-B; Figure A-B).
+In microtellite data, larger distances of a node to the root (Dnr) were strongly correlated with higher bootstrap support (??statistics??)(Figure 3 C-D; Figure 4 C-D).
+This correlation was ¿¿not??
+significant for the SNP data (??statistics??)(Figure 3 A-B; Figure 4 A-B).
+The distance of a node its precedent in direction to the root, or parent node (Dnp), explained a huge variation of in the bootstrap support for all datasets ??statistics??
+(Figure 3 E-I; Figure 4 C-D).
+In the case of the SNP data Dnp ranged from 0?
+to 0.15?, but almost all trees supported nodes that had a Dnp above 0.025?.
+The slope and R^2 of the logarithmic regression for the SNP data were very large compared to the microsatellites (??statistics??; Figure 3 E-I).
+In *P.
+cultripes*, where the number of SNPs in the dataset was larger than in *H.
+molleri*, the slope was greater (statistics??; Figure 3 E-F)
 

### 2019-07-10 code/03_phylogenies/03_subsampling_snps.R

@@ -8,101 +8,65 @@
 #DESCRIPTION: export data into structure format and create params files
 #PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
 ###.............................................................................
-#  REQUIRED FILES:
-#   Description:
-#   Inpath:
-#  OUTPUT:
-#    Description: raw genotypes plus metadata for dartseq samples
-#    Outpath:
-#  DEPENDENCIES:
-###.............................................................................
-library(adegenet)
-library(assertthat)
+
+library(magrittr)
 library(dplyr)
 library(ape)
+library(dartR)
 
-input_path1 <- paste0("data/intermediate/filt_genotypes.rds")
+#genotypes
+input_path1 <- "data/intermediate/gen_consolidated_filtered.rds"
 gen <- readRDS(file = input_path1)
-input_path2 <- paste0("data/intermediate/bt_trees.rds")
+#trees
+input_path2 <- "data/intermediate/bt_trees.rds"
 tr <- readRDS(file = input_path2)
-input_path3 <- paste0("data/intermediate/boot_support.rds")
-bt <- readRDS(file = input_path3)
-max_bt <- max(unlist(bt))
-bt <- sapply(function(x) x / max_bt)
+
+#tree metrics from datasets
+input_path3 <- "data/intermediate/bs_treemetrics.rds"
+bs <- readRDS(input_path3)
 
 source("code/parameters/boot.r")
 source("code/functions/bootstrap_trees.r")
-source("code/functions/create_trees.r")
+source("code/functions/bs_analysis.r")
 
 #subset data
-pelo <- gen$dart_pelo
-original_trees <- lapply(tr, function(x) x[[1]])
+pelo <- gen$dart_pelo %>% dartR::gi2gl()
+original_tree <- lapply(tr, function(x) x[[1]])$dart_pelo
 rm(tr)
 
 #create trees from subsampled dataset
-s.tr <- 1:length(s) %>%
-  sapply(function(j){
-  sp <- as.character(s[j])
-  temp <- list()
-  for (i in 1:n_sampling_replicas){
-    int <- sample(1:nLoc(pelo), s[j], replace = FALSE)
-    s_pelo <- pelo[, int]
[…]

### 2019-07-10 code/03_phylogenies/06_MAF_vs_phylogeny.R

@@ -8,28 +8,25 @@
 #DESCRIPTION:
 #PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
 ###.............................................................................
-
 library(adegenet)
 library(dplyr)
 library(ape)
+library(dartR)
 
-input_path1 <- "data/intermediate/bt_trees.rds"
-input_path2 <- "data/intermediate/boot_support.rds"
-tr <- readRDS(file = input_path1)
-bt <- readRDS(file = input_path2)
+input_path <- "data/intermediate/gen_consolidated_filtered.rds"
+gen <- readRDS(file = input_path)
 
-source("code/functions/bs_analysis.r")
 source("code/parameters/boot.r")
 
-#get list of dataframes with columns being:
-# dnr, distance from node to root
-# dnp, distance from node to parent node
-# bs_tm, bootstrap support
-#and rows being nodes.
+#convert from genind to genlight
+pelo <- gen[["dart_pelo"]] %>%
+  dartR::gi2gl
 
-bs_treemetrics <-
-  seq_along(tr) %>%
-    lapply(function(x) tree_metrics(tr[[x]], bt[[x]]))
-names(bs_treemetrics) <- names(tr)
+#filter pelobates genotypes accross MAF thresholds
+pelo_maf <-
+  seq_along(maf) %>%
+    lapply(function(x){
+      dartR::gl.filter.maf(pelo, threshold = maf[x], v = 5)
+      })
 
-saveRDS(bs_treemetrics, "data/intermediate/bs_treemetrics.rds")
+names(pelo_maf) <- as.character(maf)

### 2019-07-10 code/03_phylogenies/99-recycle_subsample_trees.R

@@ -1,0 +1,36 @@
+#create data frame with combined BS for subsampled trees and original ones
+s.btx <- lapply(s.bt, function(x) x[-1])
+s_bt_m <- reshape::melt(s.btx)
+s_bt_m[, 1] <- s_bt_m[, 1] / (nboot * n_sampling_replicas)
+assertthat::assert_that(max(s_bt_m[, 1]) == 1)
+s_bt_m[, 2] <- as.numeric(s_bt_m[, 2]) %>% as.factor()
+s_bt_m <- rbind(s_bt_m, reshape::melt(bt))
+names(s_bt_m) <- c("BS", "dataset")
+s_bt_m$BS <- round(s_bt_m$BS, 2)
+
+#quantiles for BS
+g <- by(data = s_bt_m$BS, INDICES = s_bt_m$dataset, quantile)
+gg <- 1:length(g) %>% sapply(function(x) g[[x]] %>% round(2))
+colnames(gg) <- names(g); rm(g)
+
+#labs for plot
+nl <- sapply(gen, adegenet::nLoc)
+lab <- c(levels(s_bt_m$dataset)[1:length(s)],
+  paste0(names(gen), rep("\n (", 4), nl, rep(")", 4)))
+
+#boxplot
+pdf(file = "data/intermediate/bs_boxplot.pdf", height = 5, width = 8)
+  boxplot(s_bt_m$BS~s_bt_m$dataset, xlab = "Number of loci",
+      ylab = "Bootstrap support",
+      main = "Change of overall boostrap support accross number of loci",
+      col = c(rep("grey", length(s)), rep(grey(0.3), 2), rep(grey(0.8), 2)),
+      pch = 20, outcol = grey(0.7), xaxt = "n", yaxt = "n")
+      axis(1, at = 1:(length(s) + 4), labels = lab, las = 2)
+      axis(2, at = seq(0, 1, 0.2), labels = seq(0, 1, 0.2), las = 1)
+      abline(v = length(s) + 0.5, lty = 2, lwd = 2)
+dev.off()
+
+
+#save trees (first tree is the real tree)
+saveRDS(s.tr, "data/intermediate/s_trees.rds")
+saveRDS(gg, "data/intermediate/s_quantiles.rds")

### 2019-07-10 code/functions/bootstrap_trees.r

@@ -52,8 +52,18 @@ midpoint_tr <- function(list_mp){
 }
 
 #get bootstrap value for nodes for rooted or unrooted trees
-get_bt <- function(list_tr){
+get_bt <- function(list_tr, external_ref_tree = NULL){
   #list_tr is a list of multiphylo objects
+  h <- names(list_tr)
+  if (!is.null(external_ref_tree)){
+    assertthat::assert_that(
+      ape::Ntip(external_ref_tree) == ape::Ntip(list_tr[[1]][1]))
+    list_tr <-
+      seq_along(list_tr) %>%
+      lapply(function(x){
+        c(external_ref_tree, list_tr[[x]])
+      })
+    }
   bt <-
     seq_along(list_tr) %>%
     sapply(function(x){
@@ -63,7 +73,7 @@ get_bt <- function(list_tr){
         pp <- ape::prop.part(list_tr[[x]][-1])
         ans <- ape::prop.clades(list_tr[[x]][[1]], part = pp, rooted = T)
       }
-      #1. for unrooted trees
+      #2. for unrooted trees
       if (all(!sapply(list_tr, ape::is.rooted))){
         phy <- reorder(list_tr[[x]][[1]], "postorder")
         ints <- phy$edge[, 2] > ape::Ntip(phy)
@@ -74,6 +84,6 @@ get_bt <- function(list_tr){
       temp[[1]] <- ans
       temp
     })
-    names(bt) <- names(list_tr)
+    names(bt) <- h
     bt
   }

### 2019-07-10 code/functions/bs_analysis.r

@@ -1,4 +1,4 @@
-tree_metrics <- function(tr, bs){
+tree_metrics <- function(tr, bs, external_ref_tree = NULL){
   #tr is a multiphylo object
   #bs is a vector with bootstrap supports
   #returns a data frame with
@@ -9,8 +9,12 @@ tree_metrics <- function(tr, bs){
   require(gridExtra)
   library(treeman)
   source("code/functions/add_bs2tm.r")
-  #
-  tr1 <- tr[1][[1]]#reference tree
+
+  if (!is.null(external_ref_tree)){
+    tr1 <- external_ref_tree
+    } else {
+      tr1 <- tr[1][[1]]
+      }#reference tree
   bs <- bs / nboot #vector with bs values normalized to 1
   #Normalize length of the tree
   # distance between all nodes (also tips) in the tree

### 2019-07-10 code/parameters/boot.r

@@ -1,6 +1,8 @@
 nboot = 1000 #number of bootstrap
 thresh = .7 #treshold for coloring nodes
 #subsample sizes for pelobates dartseq data
-s <- c(25000, 20000, 15000, 10000, 5000, 3000, 1000, 500, 200)
+s <- c(20000, 10000, 5000, 3000, 1000, 500, 200)
 #times the markers are subsampled from the original matrix of genotypes
-n_sampling_replicas <- 3
+n_sampling_replicas <- 5
+#maf thresholds
+maf <- seq(0.04, 0.1, 0.02)

### 2019-07-10 code/Plot_subsample_SNPs.R
plots subsampling

@@ -1,0 +1,90 @@
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# June 2019
+###.............................................................................
+#GOAL: Plot results from subsampling SNPs from Pelobates
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+library(dplyr)
+library(magrittr)
+
+gen <- readRDS("data/intermediate/gen_consolidated_filtered.rds")
+nloc_pelo <- gen$dart_pelo %>% adegenet::nLoc()
+
+bstree <- readRDS("data/intermediate/s_bs_treemetrics.rds") %>%
+  {c(list(
+    dart_pelo = readRDS("data/intermediate/bs_treemetrics.rds")$dart_pelo), .)}
+
+bt <- readRDS("data/intermediate/bs_treemetrics.rds")
+bt %<>% .[!grepl("dart_pelo", names(bt))]
+
+names(bstree)[names(bstree) == "dart_pelo"] <- nloc_pelo
+
+# plot parameters
+x_lab <- c("Distance to parent node", "Distance from root")
+max_dnr <- sapply(bstree, function(x) x$dnr) %>% unlist %>% max
+max_dnp <- sapply(bstree, function(x) x$dnp) %>% unlist %>% max
+x_lim <- matrix(c(0, max_dnp, 0, max_dnr), ncol = 2)
+
+#1. Point plots
+
+pdf(file = paste0("data/final/", "subsampling_bs", ".pdf"),
+    height = 9, width = 6.5)
+  par(mfrow = c(2, 1), las = 1)
+  for (i in 1:2){
+  plot(1, pch = 20,
+       xlab = x_lab[i],
+       ylab = "Bootstrap support",
+       xlim = x_lim[, i],
+       ylim = c(0, 1), type = "n")
+  for (p in seq_along(bstree)){
+    hh <- bstree[[p]]
+    points(as.numeric(as.matrix(hh[, i])),
+        hh$bs_tm, pch = 20, col = grey.colors(length(bstree))[p])
+  }
+    legend("bottomright", legend = paste(names(bstree), "loci"),
[…]

### 2019-07-10 code/03_phylogenies/06_MAF_vs_phylogeny.R
maf bootstrap

@@ -8,25 +8,88 @@
 #DESCRIPTION:
 #PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
 ###.............................................................................
-library(adegenet)
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# April 2019
+###.............................................................................
+#GOAL: analysis of bootstrap support
+#DESCRIPTION:
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
 library(dplyr)
 library(ape)
 library(dartR)
 
+#genotypes
 input_path <- "data/intermediate/gen_consolidated_filtered.rds"
 gen <- readRDS(file = input_path)
 
+#trees
+input_path2 <- "data/intermediate/bt_trees.rds"
+tr <- readRDS(file = input_path2)
+
 source("code/parameters/boot.r")
+source("code/functions/bootstrap_trees.r")
+source("code/functions/bs_analysis.r")
+
+original_tree <- lapply(tr, function(x) x[[1]])$dart_pelo
 
 #convert from genind to genlight
-pelo <- gen[["dart_pelo"]] %>%
-  dartR::gi2gl
+pelo <- gen[["dart_pelo"]]
+#remove objects within
+other(pelo) <- NULL
+#regenerate $loc.metrics to be read correctly by dartR functions.
+pelo %<>% dartR::gi2gl() %>%
+  dartR::gl.recalc.metrics()
 
 #filter pelobates genotypes accross MAF thresholds
 pelo_maf <-
   seq_along(maf) %>%
-    lapply(function(x){
-      dartR::gl.filter.maf(pelo, threshold = maf[x], v = 5)
-      })
+  lapply(function(x){
+    dartR::gl.filter.maf(pelo, threshold = maf[x], v = 5)
+  }) %>% {c(pelo,.)}
 
-names(pelo_maf) <- as.character(maf)
+all_maf <- as.character(c(0.02, maf))
+names(pelo_maf) <- all_maf
+
+#create trees from maf datasets
+nboots <- nboot / n_sampling_replicas
+
+maf.tr <-
[…]

### 2019-07-10 code/parameters/boot.r

@@ -6,3 +6,5 @@ s <- c(20000, 10000, 5000, 3000, 1000, 500, 200)
 n_sampling_replicas <- 5
 #maf thresholds
 maf <- seq(0.04, 0.1, 0.02)
+#maf sample size
+mafss <- 500

### 2019-07-10 code/Plot_maf.R
maf plots

@@ -1,0 +1,61 @@
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# June 2019
+###.............................................................................
+#GOAL: Plot results from maf filtering in SNPs from Pelobates
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+
+library(dplyr)
+library(magrittr)
+
+bstree <- readRDS("data/intermediate/maf_bs_treemetrics.rds") %>% rev
+
+# plot parameters
+x_lab <- c("Distance to parent node", "Distance from root")
+max_dnr <- sapply(bstree, function(x) x$dnr) %>% unlist %>% max
+max_dnp <- sapply(bstree, function(x) x$dnp) %>% unlist %>% max
+x_lim <- matrix(c(0, max_dnp, 0, max_dnr), ncol = 2)
+
+#1. Point plots
+
+pdf(file = paste0("data/final/", "maf_bs", ".pdf"),
+    height = 9, width = 6.5)
+par(mfrow = c(2, 1), las = 1)
+for (i in 1:2){
+  plot(1, pch = 20,
+       xlab = x_lab[i],
+       ylab = "Bootstrap support",
+       xlim = x_lim[, i],
+       ylim = c(0, 1), type = "n")
+  for (p in seq_along(bstree)){
+    hh <- bstree[[p]]
+    points(as.numeric(as.matrix(hh[, i])),
+           hh$bs_tm, pch = 20, col = grey.colors(length(bstree))[p])
+  }
+  legend("bottomright", legend = names(bstree),
+         pch = 20, col = grey.colors(length(bstree)), bty = "n")
+}
+dev.off()
+
+#2. Boxplot
+
+#dataframe where col1 is bootstrap support and col2 is dataset
+s_bt_m <-
+  bstree %>% lapply(function(x) dplyr::select(x, bs_tm)) %>%
+  reshape2::melt() %>%
+  dplyr::select(-variable) %>%
+  dplyr::rename(dataset = 2) %>%
+  tibble::as_tibble() %>%
+  dplyr::mutate(dataset = as.factor(dataset))
+
+#boxplot
+pdf(file = "data/final/maf_bs_boxplot.pdf", height = 5.5, width = 8)
+par(las = 1)
[…]

### 2019-07-10 code/rmarkdown/Figures.rmd
map ancestries

@@
 In each tip, pie chart depicts the ancestry proportions from CLUMPAK consensus solutions for the major cluster for K = 4 together with the population ID from Table 1.
+
+##Figure 5
+
+??*to decide between top or bottom panel*??
+*Top*.
+Median (horizontal lines) and 25% and 75% quantiles (grey boxes) of the bootstrap support for all nodes in the tree (top panel) from the original datasets and from the subsampled loci.
+*Bottom*.
+Bootstrap support for trees from subsampled SNPs from *P.
+cultripes* (blue dots) and fitted values (solid line) from a linear model on the natural logarithm of the number of loci (R^2 = 0.34; p-value = < 0.001).
+
+##Figure 6
+
+??why are there transparent pies on K8. check code and export as bitmap??
+Spatial distribution of localities sampled with pies representing averaged proportion of inferred ancestries from K = 2 to K = 8, for SNP and microsatellite data, in STRUCTURE.
+The shaded area represent the distributions of *H.
+molleri* and *P.
+cultripes*.

### 2019-07-10 code/rmarkdown/article.Rmd
map ancestries

@@
 We used thresholds for allele balance of 0.15 and 0.85, and removed loci for which their coverage was 3.5 times higher than the median coverage (O'Leary et al.
-2018), after which, we filtered loci with a call rate (proportion of samples with a call) below 0.8, retained only one SNP per contig, a set minor allele frequency threshold of 0.02 and removed loci whose alleles deviated from Hardy-Weinberg proportions(O'Leary et al.
+2018), after which, we filtered loci with a call rate (proportion of samples with a call) below 0.8, retained only one SNP per contig, a set minor allele frequency threshold (MAF) of 0.02 and removed loci whose alleles deviated from Hardy-Weinberg proportions(O'Leary et al.
 2018).
@@
 ## 2.4 Phylogenetic structure
-
 We evaluated the phylogenetic relationships between the populations using the neighbor-joining (NJ) algorithm (Saitou and Nei 1987).
@@
 *Number of markers and bootstrap support*.
+The number of SNPs can affect the phylogeographic patterns inferred from RAD-seq data (Hodel et al.
+2017).
+We evaluated the effect of the number of SNPs on the support of the tree reconstructions.
+We sampled random subsets of 20000, 10000, 5000, 3000, 1000, 500 and 200 loci from the SNP dataset of *P.
+cultripes*.
+This process was repeated 5 times to overcome possible biases on marker choice.
+On each of these 5 subsets we computed 200 bootstrapped trees as above.
+We calculated the bootstrap supports by counting the number of partitions of the NJ tree from the full dataset (see above, `r adegenet::nLoc(gen$dart_pelo)` loci) the were present in the 1000 bootstrapped trees.
+The number of loci was linearized applying a natural logarithm and its relation with the bootstrap support was evaluated with a linear model in R.
 
 *MAF and bootstrap support*.
+Loci can contain different amount of information.
+A way to quantify this information is by the MAF (ref).
+MAF thresholds used in filtering steps can greatly affect population genetic inferences, since allele frequencies are related to the mutation rate of their locus, thus  yielding information about different evolutionary scales (Linck and Battey 2019).
+Indeed, singletons blur the results of model-based genetic structure analysis (Linck and Battey 2019) and stringent MAF threshold (ie.
+0.10) seem to yield best results (Haasl and Payseur 2011).
+However, their possible effect on the evolutionary relationships between populations has not been studied.
+We generated SNP datasets by applying MAF thresholds of 0.02, 0.04, 0.06, 0.08 and 0.10, to the SNPs from *P.
+cultripes*.
+From each dataset we sampled 500 random loci 5 times and computed bootstrap support from the resulting trees as above.
+We chose 500 loci because for this particular dataset it is a number of loci sensitive to changes on the overall bootstrap support of the tree (see Results).
+The relation between the bootstrap supports and MAF filters were assessed with a linear model in R.
 
@@
 After that, the distinct marker sets showed inconsistent trends (Figure 2I).
-*clumpak table*
+??*clumpak table* talk about major/minor clusters in clumpak or similarities betweend ancestry matrices between mean population ancestries for usat/snp data or between n solutions for K??
+For K = 2, the genetic structure was similar for microtellite and SNP data from both species (Figure 6), differentiating a norther and southern lineage for *H.
+molleri*, and a central-western and north-eastern linage for *P.
+cultripes*.
+However, from K = 3 to K = 8, the spatial patterns of genetic structure for both species showed moderate disagreements between SNPs versus microsatellites (Figure 6), which increased with higher values of K.
+These differences were more marked in *H.
+molleri*.
+For instance, in *H.
+molleri* the southeastern population of Ojos de Villaverde stood out as very distinct from K = 4 from SNPs whereas microtellites detected this singularity at K = 5.??should I focus on the biological patterns to allow a discussion on the phylogegrphy for these species or just on the marker comparison??
 
@@
 However, the topology of the trees for microtellites were little supported.
-Furthermore, it neither provided a solid grouping of the samples from the same populations and it was not able to explain the relationships at different depths in the trees (Figure 4 C-D).
+Furthermore, it neither provided a solid grouping of the samples from the same populations and it was not able to explain the relationships at different depths in the trees (Figure 4 C-D).??look at where the pops are located and talk about the spatial coherence of the phylogegraphic patterns.
+what can this info provides us regarding the phylogeography of these species respect to microtellites??
 
@@
 *Number of markers and bootstrap support*.
+The bootstrap support was highest for the dataset with most loci (n = 33144), for which most nodes had a support equal to 1 (Figure 5).
+The bootstrap support decreased exponentially as the number of markers decreased down to 200 (Figure 5).
+The relation between bootstrap support and natural logarithm of number of loci was well explained by a linear model (R^2 = 0.34; p-value = < 0.001).
+The overall BS for the tree from the SNP dataset with lowest number of loci (n = 200 loci) was comparable to the BS for tree from the microtellite data of *H.
+molleri* (Figure 5).
 
-*MAF and bootstrap support*.
+*MAF and bootstrap support*. There was no effect of the MAF thresholds on the bootstrap support from the trees (t-test = 0.68, p-value = 0.5; Figure S4)
 
@@
 
-* Overview
-
-* Key result 1
-
-* Key result 2
-The little bootstrap support we describe in phylogenetic trees from microtellite data is not surprising, given that often several tens of microsatellites or more are needed to recover the correct topology of trees (Nei et al.
-1983; Takezaki and Nei 1996)
-
-* Key result 3
-
-* Concluding paragraph
-
 # AUTHORS' CONTRIBUTIONS
-
-TODO
 

### 2019-07-10 code/Plot_maf.R

@@ -59,3 +59,21 @@ boxplot(s_bt_m$value~s_bt_m$dataset, xlab = "MAF threshold",
         pch = 20, outcol = grey(0.7))
 
 dev.off()
+
+#3. Point plot linear model
+h <- s_bt_m
+# model
+m1 <- lm(value ~ as.numeric(as.character(dataset)), data = h)
+summary(m1)
+pr <- predict(m1,
+  newdata = data.frame(dataset = seq(0, 0.1, length.out = 1000)))
+
+pdf(file = "data/final/maf_model.pdf", height = 4.5, width = 6)
+plot(seq(0, 0.1, length.out = 1000),
+     pr,
+     ylab = "Bootstrap support",
+     xlab = "MAF", las = 1, type = "l", ylim = c(0, 1))
+points(jitter(as.numeric(as.character(h$dataset))),
+       h$value,
+       col = adjustcolor("blue", alpha.f = 0.2), pch = 20)
+dev.off()

### 2019-07-10 code/Plot_subsample_SNPs.R

@@ -88,3 +88,20 @@ axis(1, at = seq_along(levels(s_bt_m$dataset)), labels = lab,
 axis(2, at = seq(0, 1, 0.2), labels = seq(0, 1, 0.2), las = 1, cex.axis = 0.8)
 abline(v = length(bstree) - 0.5, lty = 2, lwd = 2)
 dev.off()
+
+#3. Point plot linear model
+meanbs <- sapply(bstree, function(x) mean(x$bs_tm))
+h <- s_bt_m %>% filter(!grepl("^[a-z]", dataset))
+# model
+m1 <- lm(value ~ log(as.numeric(as.character(dataset))), data = h)
+summary(m1)
+pr <- predict(m1, newdata = data.frame(dataset = seq(100, 33000, 1)))
+
+pdf(file = "data/final/no_loci_model.pdf", height = 4.5, width = 6)
+plot(pr,
+     ylab = "Bootstrap support",
+     xlab = "Number of loci", las = 1, type = "l", ylim = c(0, 1))
+points(jitter(as.numeric(as.character(h$dataset))),
+       h$value,
+       col = adjustcolor("blue", alpha.f = 0.2), pch = 20)
+dev.off()

### 2019-07-10 code/parameters/structure_colors.r

@@ -1,0 +1,2 @@
+str_col <- c("#0099e5ff", "#ff6600ff", "#33004cff", "#0d660dff",
+             "#99194cff", "#feb1b1ff", "#fef21dff")

### 2019-07-10 code/plot_map_ancestry_population.R

@@ -1,0 +1,111 @@
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# June 2019
+###.............................................................................
+#GOAL: Plot map with ancestries per population
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+
+library(dplyr)
+library(magrittr)
+library(sf)
+library(raster)
+library(adegenet)
+
+#vector to reorder datasets
+dor <- c("dart_hyla", "usat_hyla", "dart_pelo", "usat_pelo")
+
+#genotypes
+gen <- readRDS("data/intermediate/gen_consolidated_filtered.rds")
+gen <- gen[match(dor, names(gen))]
+
+#ancestries for major clusters
+qclumpak <- readRDS("data/intermediate/clumpak_major.rds")
+# remove subset of loci
+qclumpak %<>% .[grep("^[a-z]", names(qclumpak))] %>% {.[match(dor, names(.))]}
+
+#load plotting function
+source("code/functions/map_distribution.r")
+#load functions
+source("code/functions/centroid_sf.r")
+#colors for ancestries
+source("code/parameters/structure_colors.r")
+
+#assertions
+
+assertthat::assert_that(seq_along(gen) %>% sapply(function(x){
+  names(gen[[1]]@other$metadata) == c("sample_id", "locality",  "geometry")
+}) %>% all())
+
+#1. Read raster and shape files
+#raster
+p_raster <- "data/raw/raster_iberian_peninsula.grd"
+mapr <- raster::raster(p_raster)
+#IUCN distribution
+p_pelo <- "data/raw/redlist_species_data_44f009a9-b3b2-4608-a818-24c3dd714900"
+p_hyla <- "data/raw/redlist_species_data_eff4cf11-63bb-4131-bba3-104c3358697b"
+shp_pelo <- rgdal::readOGR(dsn = p_pelo, layer = "data_0") %>%
+  sp::spTransform(mapr@crs)
+shp_hyla <- rgdal::readOGR(dsn = p_hyla, layer = "data_0") %>%
+  sp::spTransform(mapr@crs)
+#check I loaded the right layers
[…]

### 2019-07-16 code/rmarkdown/article.Rmd
new hyla data. sample codes corrected

@@
 Number of equations: xx  \newline
-Keywords: microtellites, DArTseq, SNPs, neighbour joining, structure, phylogeography, Hyla molleri, Pelobates cultripes, Iberian Peninsula   \newline
+Keywords: microtellites, DArTseq, SNPs, neighbour joining, structure, phylogeography, *Hyla molleri*, *Pelobates cultripes*, Iberian Peninsula   \newline
 Spelling: British English (en-GB)   \newline
@@
 On each of these 5 subsets we computed 200 bootstrapped trees as above.
-We calculated the bootstrap supports by counting the number of partitions of the NJ tree from the full dataset (see above, `r adegenet::nLoc(gen$dart_pelo)` loci) the were present in the 1000 bootstrapped trees.
+We calculated the bootstrap supports by counting the number of partitions of the NJ tree from the full dataset (see above, 33144 loci) the were present in the 1000 bootstrapped trees.
 The number of loci was linearized applying a natural logarithm and its relation with the bootstrap support was evaluated with a linear model in R.

### 2019-07-16 code/01_prepare_data/01_read_data.R

@@ -19,16 +19,22 @@
 
 library(dartR)
 library(dplyr)
+library(magrittr)
 
 #1. DArTseq
 #path to singlerow dartseq SNPs genotypes
 path_2_genotypes <- c(
-  "data/raw/Report_DHyl17-2984_1_moreOrders_SNP_singlerow_2.csv",
+  "data/raw/Report_DHyl17-2984_SNP_mapping_2.csv",
   "data/raw/Report_DPelo19-4013_SNP_mapping_2.csv")
 
 #read singlerow dartseq file into genlight object.
 gen_dart <- lapply(path_2_genotypes, dartR::gl.read.dart)
 names(gen_dart) <- c("hyla", "pelo")
+
+#remove replica from Pelobates:
+    #for Pelobates Ana included a blind
+    #replica named 108541 from sample 106854.
+gen_dart$pelo <- gl.drop.ind(gen_dart$pelo, "108541")
 #read metadata
 source("code/functions/read_metadata.r")
 gen_dart %<>% metadata_Dartseq_Hyla() %>% metadata_Dartseq_Pelobates()

### 2019-07-16 code/01_prepare_data/02_genotype_filtering.R

@@ -21,7 +21,6 @@ library(magrittr)
 
 source("code/functions/gl_stats.r")
 source("code/functions/ind_miss.r")
-source("code/functions/remove_replica.r")
 source("code/functions/coverage_filt.r")
 source("code/parameters/gl_cleaning.r")
 
@@ -34,11 +33,8 @@ raw_stats <- geno_stats(genlist = gen, data = "raw")
 ##################
 #Filtering
 sink(file = "data/intermediate/filtering.log")
-  #remove replica from Pelobates:
-      #for Pelobates Ana included a blind
-      #replica named 108541 from sample 106854.
-gen$dart_pelo <- remove_replica(gen = gen$dart_pelo, "108541", "106854")
-  #filter SNPs
+
+#filter SNPs
 
 dart_filt <- grep("dart", names(gen)) %>% #only for dart genotypes
   {genfiltnames <<- names(gen)[.]; .} %>%

### 2019-07-16 code/01_prepare_data/04_Filter_populations.R

@@ -26,13 +26,13 @@ source("code/functions/map_distribution.r")
 
 #Plot samples in map so as to decide which populations to keep
 plot_samples(meta = gen,
-  name_plot = "data/intermediate/pre_sample_fil_distribution.pdf")
+  name_plot = "data/intermediate/pre_sample_fil_distribution.pdf", res = 5000)
 #Create list with populations to remove
 #based on plotting samples on a map, I removed the following populations
 usat_hyla_remove_pop <- c("Segonnes", "Larués", "Renales", "Torrecaballeros",
-  "Valdemanco", "Colmenar Viejo", "Peñalara", "Albires", "Río Montoro",
+  "Valdemanco", "Colmenar Viejo", "Peñalara", "Río Montoro",
   "Albires", "Serra de Montejunto", "Mindelo", "El Puerto", "Valdeazores",
-  "Valle de Santiago", "Boticas", "Fuenterrebollo")
+  "Valle de Santiago", "Boticas", "Fuenterrebollo", "Ciudad Rodrigo")
 dart_hyla_remove_pop <- c() #none filtered
 usat_pelo_remove_pop <- c("Valflaunes", "Valdunciel", "Porto Covo",
   "Cabezarrubias del Puerto")
@@ -76,12 +76,12 @@ gen$usat_hyla$other$metadata$locality <-
   gsub("Castelo de Vide", "Beira", gen$usat_hyla$other$metadata$locality)
 # merge populations
 #  1
-gen$dart_hyla$other$metadata$locality %<>%
-  gsub(pattern = "Peña de Francia",
-    replacement = "Peña de Francia - Ciudad Rodrigo")
-gen$usat_hyla$other$metadata$locality %<>%
-  gsub(pattern = "Ciudad Rodrigo",
-    replacement = "Peña de Francia - Ciudad Rodrigo")
+# gen$dart_hyla$other$metadata$locality %<>%
+#   gsub(pattern = "Peña de Francia",
+#     replacement = "Peña de Francia - Ciudad Rodrigo")
+# gen$usat_hyla$other$metadata$locality %<>%
+#   gsub(pattern = "Ciudad Rodrigo",
+#     replacement = "Peña de Francia - Ciudad Rodrigo")
 #   2
 gen$dart_hyla$other$metadata$locality %<>%
   gsub(pattern = "Saceruela|Río Esteras",
@@ -136,6 +136,7 @@ gen %<>% lapply(function(x){
[…]

### 2019-07-16 code/functions/read_metadata.r

@@ -34,14 +34,16 @@ assertthat::assert_that(!is.null(c(m[["dartseqID"]],
 }
 
 metadata_Dartseq_Pelobates <- function(x){
-  f <- "data/raw/2019-01-03_meta_Pelobates.csv"
-  m <- read.csv(f)
-  m %<>% mutate(locality = gsub(pattern = ",.*$", replacement = "", site)) %>%
+  f <- "data/raw/meta_pelobates.csv"
+  m <- read.delim(f, sep = ";")
+  m %<>%
+    mutate(locality = gsub(pattern = ",.*$", replacement = "", site)) %>%
+    mutate(collector = gsub(pattern = " ", replacement = "", collector)) %>%
     rename(full_locality = site) %>%
     rename(dartseqID = vaucher) %>%
     rename(latitude = lat) %>%
     rename(longitude = lon) %>% dplyr::as_tibble() %>%
-    {sp::SpatialPointsDataFrame(data = .[, c(1, 6), drop = FALSE],
+    {sp::SpatialPointsDataFrame(data = .[, c(1, 6, 7), drop = FALSE],
                                 coords = .[, c("longitude", "latitude")])} %>%
     as("sf") %>%
     sf::st_set_crs(4326)
@@ -52,12 +54,20 @@ metadata_Dartseq_Pelobates <- function(x){
   assertthat::assert_that(all(g_dartseq_id %in% m_darseq_id), msg =
   "some names in genotypes are missing in metadata")
   #fix pop names
-  assertthat::assert_that(!is.null(c(m[["dartseqID"]], m[["locality"]])))
+  assertthat::assert_that(!is.null(c(m[["collector"]], m[["locality"]])))
+
   pop(x$pelo) <- sapply(g_dartseq_id, function(x){
     which(m[["dartseqID"]] == x) %>% m[["locality"]][.]
   }) %>% unlist()
+
+  indNames(x$pelo) <- sapply(g_dartseq_id, function(x){
+    which(m[["dartseqID"]] == x) %>% m[["collector"]][.]
+  }) %>% unlist()
+
+  attr(x = pop(x$pelo), "names") <- indNames(x$pelo)
+  m %<>% dplyr::select(-dartseqID)
   metadata_dartpelo <<- m %>% #select only samples present in the genlight object
-                filter(dartseqID %in% adegenet::indNames(x$pelo)) %>%
-                rename(sample_id = dartseqID)
[…]

### 2019-07-16 code/functions/read_usats.r

@@ -103,6 +103,11 @@ read_Hyla_molleri <- function() {
        dplyr::rename(sample_id = V1, locality = V2)
        g[["sample_id"]] <- gsub(pattern = "^.*#", #m
                       replacement = "", g[["sample_id"]]) #m
+  #replace museum codes with collector codes
+  g$sample_id %<>%
+    gsub(pattern = "MNCN11068", replacement = "HaGa1") %>%
+    gsub(pattern = "MNCN11078", replacement = "HaMj1") %>%
+    gsub(pattern = "MNCN11086", replacement = "HaT3")
 
   s <- read.table("data/raw/table-1-data.csv",
               header = TRUE, sep = ",", stringsAsFactors = FALSE) %>%

### 2019-07-16 code/02_structure/01_Prepare_structureK1.R
subsampling hyla

@@ -8,14 +8,7 @@
 #DESCRIPTION: export data into structure format and create params files
 #PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
 ###.............................................................................
-#  REQUIRED FILES:
-#   Description:
-#   Inpath:
-#  OUTPUT:
-#    Description: raw genotypes plus metadata for dartseq samples
-#    Outpath:
-#  DEPENDENCIES:
-###.............................................................................
+
 library(dartR)
 library(dplyr)
 
@@ -25,12 +18,3 @@ source("code/functions/edit_params.r")
 
 #1. Create STRUCTURE input files from dart and usat genotypes
 create_str_input(mode = "K1")
-
-#2. Create STRUCTURE input files from SNP downsampling genotypes
-  #genotypes for pelobates
-pelo <- gen$dart_pelo
-  #load subsampling sizes
-source("code/parameters/boot.r")
-s <- s[!(s %in% c(15000, 25000))]#edit subsampling vector
-
-create_str_input_subsets(mode = "K1")

### 2019-07-16 code/03_phylogenies/03_subsampling_snps.R

@@ -16,7 +16,8 @@ library(dartR)
 
 #genotypes
 input_path1 <- "data/intermediate/gen_consolidated_filtered.rds"
-gen <- readRDS(file = input_path1)
+gen <- readRDS(file = input_path1) %>%
+  {.[grepl("dart", names(.))]} #only select dart datasets
 #trees
 input_path2 <- "data/intermediate/bt_trees.rds"
 tr <- readRDS(file = input_path2)
@@ -29,44 +30,62 @@ source("code/parameters/boot.r")
 source("code/functions/bootstrap_trees.r")
 source("code/functions/bs_analysis.r")
 
-#subset data
-pelo <- gen$dart_pelo %>% dartR::gi2gl()
-original_tree <- lapply(tr, function(x) x[[1]])$dart_pelo
-rm(tr)
+#empty list
+out_trees <- list()
 
-#create trees from subsampled dataset
-nboots <- nboot / n_sampling_replicas
-
-s.tr <-
-  seq_along(s) %>%
-  sapply(function(j){
-    sp <- as.character(s[j])
-    temp <- list()
-    for (i in 1:n_sampling_replicas){
-      int <- sample(1:nLoc(pelo), s[j], replace = FALSE)
-      s_pelo <- pelo[, int]
-      if (i == 1)
-        boots_trees <- bootstrap_nj(s_pelo, nboots, type = "snp")
-      if (i > 1)
-        boots_trees <- c(boots_trees, bootstrap_nj(s_pelo, nboots, type = "snp"))
-    }
-    temp[[1]] <- boots_trees
-    names(temp) <- sp
-    temp
-  })
-
-#root trees
-tr_rooted <- midpoint_tr(s.tr)
-
-#get bootstrap#the function internally assigns 0's to all NA values in
-#bootstrap calculations.
-bt <- get_bt(tr_rooted, external_ref_tree = original_tree)
-
+#calculate bootstrap support with tree metrics (dnr and dnp) for trees made from
+#datasets which are a subsample of the loci from larger matrices.
 bs_treemetrics.s <-
-  seq_along(tr_rooted) %>%
-  lapply(function(x) tree_metrics(tr[[x]], bt[[x]],
-    external_ref_tree = original_tree))
-names(bs_treemetrics.s) <- names(tr_rooted)
+seq_along(gen) %>%
+  lapply(function(z){
+#   subset data
[…]

### 2019-07-17 code/03_phylogenies/06_MAF_vs_phylogeny.R
maf subsampling hyla

@@ -8,23 +8,15 @@
 #DESCRIPTION:
 #PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
 ###.............................................................................
-###.............................................................................
-# (c) Miguel Camacho Sánchez
-# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
-# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
-# April 2019
-###.............................................................................
-#GOAL: analysis of bootstrap support
-#DESCRIPTION:
-#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
-###.............................................................................
+
 library(dplyr)
 library(ape)
 library(dartR)
 
 #genotypes
 input_path <- "data/intermediate/gen_consolidated_filtered.rds"
-gen <- readRDS(file = input_path)
+gen <- readRDS(file = input_path) %>%
+  {.[grepl("dart", names(.))]} #only select dart datasets
 
 #trees
 input_path2 <- "data/intermediate/bt_trees.rds"
@@ -34,62 +26,80 @@ source("code/parameters/boot.r")
 source("code/functions/bootstrap_trees.r")
 source("code/functions/bs_analysis.r")
 
-original_tree <- lapply(tr, function(x) x[[1]])$dart_pelo
+#empty list
+out_trees <- list()
 
-#convert from genind to genlight
-pelo <- gen[["dart_pelo"]]
-#remove objects within
-other(pelo) <- NULL
-#regenerate $loc.metrics to be read correctly by dartR functions.
-pelo %<>% dartR::gi2gl() %>%
-  dartR::gl.recalc.metrics()
+#calculate bootstrap support across different mafs
 
-#filter pelobates genotypes accross MAF thresholds
-pelo_maf <-
-  seq_along(maf) %>%
-  lapply(function(x){
-    dartR::gl.filter.maf(pelo, threshold = maf[x], v = 5)
-  }) %>% {c(pelo,.)}
-
-all_maf <- as.character(c(0.02, maf))
-names(pelo_maf) <- all_maf
-
-#create trees from maf datasets
-nboots <- nboot / n_sampling_replicas
-
-maf.tr <-
-  seq_along(all_maf) %>%
-  sapply(function(j){
-    sp <- all_maf[j]
[…]

### 2019-07-17 code/02_structure/03_Read_StructureK1.R
create structure subsampling hyla

@@ -1,4 +1,14 @@
-library(ggplot2)
+###.............................................................................
+# (c) Miguel Camacho Sánchez
+# miguelcamachosanchez@gmail.com // miguelcamachosanchez.weebly.com
+# https://scholar.google.co.uk/citations?user=1M02-S4AAAAJ&hl=en
+# July 2019
+###.............................................................................
+#GOAL: get lambda and check convergence
+#DESCRIPTION:
+#PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
+###.............................................................................
+
 #check convergence for all runs
 #dataset#k
 source("code/functions/read_lk_str.r")
@@ -12,16 +22,17 @@ lambda <- get_lambda()
   pcol <- nlevels(p$V1) %>% sqrt %>% floor()
   it <- ncol(p) - 2
 
-for (datset in 1:nlevels(p$V1)){#for each dataset
-  levelp <- levels(p$V1)[datset] #name of dataset
-  pd <- filter(p, V1 == levelp)#filter dataset
+#plot convergence
+for (datset in 1:nlevels(p$dataset)){#for each dataset
+  levelp <- levels(p$dataset)[datset] #name of dataset
+  pd <- filter(p, dataset == levelp)#filter dataset
   pdf(paste0("data/final/convergence", levelp, ".pdf"))
-  par(mfrow = n2mfrow(length(unique(pd$V2))))
-  for (k in min(pd$V2):max(pd$V2)){
-    pdk <- filter(pd, V2 == k)
+  par(mfrow = n2mfrow(length(unique(pd$k))))
+  for (K in min(pd$k):max(pd$k)){
+    pdk <- filter(pd, K == k)
     y_lim <- range(pdk[, -c(1, 2)])
     plot(1, type = "n", xlim = c(1, it), ylim = y_lim,
-         main = paste0("K = ", k), ylab = "Ln Like",
+         main = paste0("K = ", K), ylab = "Ln Like",
          xlab = "iterations x UPDATEFREQ (50) ")
     for (lin in 1:nrow(pdk)){
       lines(x = c(1:it),  y = pdk[lin, -c(1, 2)])
@@ -29,32 +40,5 @@ for (datset in 1:nlevels(p$V1)){#for each dataset
   }
   dev.off()
 }
-#check for convergence
-#(c) Jeff Hanson
-#issue: read from parameters
[…]

### 2019-07-17 code/02_structure/04_Prepare_structure_allK.R

@@ -8,29 +8,19 @@
 #DESCRIPTION: export data into structure format and create params files
 #PROJECT: usat_snp (https://github.com/csmiguel/usat_snp)
 ###.............................................................................
-#  REQUIRED FILES:
-#   Description:
-#   Inpath:
-#  OUTPUT:
-#    Description: raw genotypes plus metadata for dartseq samples
-#    Outpath:
-#  DEPENDENCIES:
-###.............................................................................
+
 library(dartR)
 library(dplyr)
 
 input_path <- "data/intermediate/gen_consolidated_filtered.rds"
 gen <- readRDS(file = input_path)
 source("code/functions/edit_params.r")
+source("code/parameters/boot.r")
 
 #1. Create STRUCTURE input files from dart and usat genotypes
 create_str_input(mode = "normal")
 
 #2. Create STRUCTURE input files from SNP downsampling genotypes
-  #genotypes for pelobates
-pelo <- gen$dart_pelo
-  #load subsampling sizes
-source("code/parameters/boot.r")
-s <- s[!(s %in% c(15000, 25000))]#edit vector with subsampling
 
-create_str_input_subsets(mode = "normal")
+create_str_input_subsets(sp = "hyla")
+create_str_input_subsets(sp = "pelo")

### 2019-07-17 code/functions/edit_params.r

@@ -34,7 +34,7 @@ edit_mainparams <- function(gl = gl, name = name,
 
 #create extraparams file for structure input
 edit_extraparams <- function(gl = gl, name = name,
-  str_file = str_file, dir_str = dir_str, mode = mode){
+  str_file = str_file, dir_str = dir_str, mode = mode, lambdapar = lambdapar){
   param <- readLines("data/raw/extraparams")
   hh <- grep("#define UPDATEFREQ ", param)
   if (grepl("dart*", name) | is.numeric(name)){
@@ -49,6 +49,10 @@ edit_extraparams <- function(gl = gl, name = name,
   } else if (mode == "normal"){
     param[hh] <- sub("xx", 0, param[hh])
   }
+  hh <- grep("#define LAMBDA ", param)
+  if (mode == "normal"){
+    param[hh] <- sub("1", lambdapar, param[hh])
+  }
   assertthat::assert_that(all(grepl("xx", param) == F),
     msg = "wrong replacement in extraparams file")
   mainparams_path <- file.path(dir_str, "extraparams")
@@ -61,13 +65,21 @@ create_str_input <- function(mode = c("normal", "K1")){
 seq_along(gen) %>%
   sapply(function(x){
     if (mode == "K1") dirn <- "strK1_" else if (mode == "normal") dirn <- "str_"
+    #determine lambda
+    assertthat::assert_that(!is.null(lambda))
+    if (grepl("usat", names(gen)[x])){
+      l <- 1
+      } else if (grepl("dart", names(gen)[x])){
+        l <- lambda[grepl(names(gen)[x], lambda$dataset), 2]
+      }
+    #end determine lambda
     dir_str <- file.path("data/intermediate", paste0(dirn, names(gen)[x]))
     dir.create(dir_str, showWarnings = T)
     str_file <- file.path(dir_str, paste0(names(gen)[x], ".str"))
     edit_mainparams(gl = gen[[x]], name = names(gen)[x],
       str_file = str_file, dir_str = dir_str)
     edit_extraparams(gl = gen[[x]], name = names(gen)[x],
-      str_file = str_file, dir_str = dir_str, mode = mode)
+      str_file = str_file, dir_str = dir_str, mode = mode, lambdapar = l)
     if (class(gen[[x]]) == "genind"){
[…]

### 2019-07-17 code/03_phylogenies/99_04_model_fitting.R
rename

@@ -10,10 +10,9 @@
 ###.............................................................................
 library(dplyr)
 
-bs_treemetrics <- readRDS <- ("data/intermediate/bs_treemetrics.rds")
+bs_treemetrics <- readRDS("data/intermediate/bs_treemetrics.rds")
 #i tried to fit glm with binomial distirbution and others but they do not fit well the data.
-m1 <- lm(bs_tm~log(dnp), data = bs_treemetrics$dart_pelo)
-m2 <- glm(bs_tm~dnp*dnr, data = bs_treemetrics$dart_pelo, family = "binomial")
+m1 <- glm(bs_tm ~ dnp + dnr, data = bs_treemetrics$dart_pelo, family = "binomial")
 
 summary(m1)
 plot(bs_tm~dnp, data = bs_treemetrics$dart_pelo)

### 2019-07-17 code/FigureS1_map_samples.R

@@ -13,4 +13,4 @@ source("code/functions/map_distribution.r")
 gen <- readRDS("data/intermediate/gen_consolidated_filtered.rds")
 
 plot_samples(meta = gen,
-  name_plot = "data/final/Figure1_map_samples.pdf", res = 50000)
+  name_plot = "data/final/FigureS1_map_samples.pdf", res = 50000)

### 2019-07-17 code/Figure4_markertype-bs.R
plot map + BS

@@ -41,3 +41,48 @@ for (i in 1:2){
       }
   dev.off()
 }
+
+#plot 2
+plot_data <-
+seq_along(bstree) %>%
+  lapply(function(x){
+  bstree[[x]] %>% mutate(dataset = as.factor(names(bstree)[x]))
+    }) %>% {do.call(rbind, .)} %>%
+  dplyr::mutate(species = as.character(dataset) %>%
+                stringr::str_remove("[aA-zA]++ ")) %>%
+  dplyr::mutate(marker = as.character(dataset) %>%
+                stringr::str_remove(" .*$")) %>%
+  {reshape2::melt(., id.vars = names(.)[-grep("dnr|dnp", names(.))])} %>%
+    tibble::as_tibble() %>%
+  dplyr::select(-dataset) %>%
+  dplyr::rename(dist = variable)
+
+library(cowplot)
+library(ggplot2)
+
+labels_distance <- c(dnp = "Dnp", dnr = "Dnr")
+p <-
+ggplot(data = plot_data) +
+  geom_point(mapping = aes(x = value, y = bs_tm,
+    colour = marker, shape = marker)) +
+  facet_grid(species ~ dist, scales = "free_x",
+             labeller = labeller(dist = labels_distance)) +
+  labs(y = "Bootstrap support", x = "") +
+  theme(
+    strip.text.y = element_text(size = 10, face = "italic"),
+    strip.text.x = element_text(size = 10)
+  ) +
+  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
+  scale_shape_manual(values = c(5, 16)) +
+  theme(legend.position = c(0.35, 0.3),
+        legend.background = element_rect(size = 0.2, linetype = "solid",
+                                 colour = "black"),
+        legend.text = element_text(colour = "black", size = 10),
+        legend.title =  element_blank())
+
+p2 <- add_sub(p, label = c("Dnp"), x = 0.25, y = 3, vjust = 1, size = 10,
+  vpadding = grid::unit(0, "lines"))
+p3 <- add_sub(p2, label = c("Dnr"), x = 0.75, y = 3, vjust = 0, size = 10,
+  vpadding = grid::unit(0, "lines"))
+ggdraw(p3)
+ggsave("data/final/Figure4_ggplot.pdf", height = 6, width = 8)

### 2019-07-17 code/FigureS1_map_samples.R

@@ -13,4 +13,4 @@ source("code/functions/map_distribution.r")
 gen <- readRDS("data/intermediate/gen_consolidated_filtered.rds")
 
 plot_samples(meta = gen,
-  name_plot = "data/final/FigureS1_map_samples.pdf", res = 50000)
+  name_plot = "data/final/FigureS1_map_samples.pdf", res = 50000, mode = 2)

### 2019-07-17 code/functions/map_distribution.r

@@ -42,6 +42,7 @@ micro_col <- "blue"
 snp_col <- "red"
 pch_snp <- 20
 pch_usat <- 2
+cex_pt <- c(1, 1.4)
 
 pdf(outfile, width = size * ratio, height = size)
 par(mfrow = c(1, 2))
@@ -59,14 +60,14 @@ for (i in seq_along(x)){
        axes = T, box = FALSE, ylab = "Latitude", xlab = "Longitude")
   plot(get(ls()[grepl(paste0("shp_", x[i]), ls())]),  #IUCN distribution
        add = T, border = "black", col = grey(0.5, alpha = 0.3))
-  plot(sf::st_geometry(centroids_dart), add = T, pch = pch_snp, cex = 1.4,
+  plot(sf::st_geometry(centroids_dart), add = T, pch = pch_snp, cex = cex_pt[2],
        col = snp_col) #snps samples
-  plot(sf::st_geometry(centroids_usat), add = T, pch = pch_usat, cex = 1,
+  plot(sf::st_geometry(centroids_usat), add = T, pch = pch_usat, cex = cex_pt[1],
        col = micro_col) #usats samples
   title(y[i], font.main = 3, cex.main = 0.8)
   if (i == 1){
     legend(y = extent(mapr)@ymax, x = extent(mapr)@xmin,
-      legend = c("Microsatellites", "SNPs"), cex = 0.7,
+      legend = c("Microsatellites", "SNPs"), pt.cex = cex_pt, cex = 0.7,
       pch = c(pch_usat, pch_snp), bty = "n", col = c(micro_col, snp_col))
   }
   off <- 0.5
